---
title: "Step 1: Sample QC [3 batches]"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

filepathp = '/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output/'
filepaths = '/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output/'
#setwd(filepaths)
setwd("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/qc/workflowr/Serreze-T1D_Workflow")

#libraries
library(broman)
library(qtl2)
library(qtlcharts)
library(ggplot2)
library(ggrepel)
#library(DOQTL)
library(mclust)
#source("code/reconst_utils.R")
library("kableExtra")
library("knitr")
library("fst")
library(data.table) 
library(tidyr)     
library(mclust)     
library(rhdf5)      
library(optparse)
library(dplyr)
library(cluster)
library(rhdf5)
library(ggplot2)
library(readxl) 

```
This script is running genotype QC on raw data (with some outcomes already seen in the [project at a glace](summary.html)). Here, we first load the R/qtl2 package and the data.
We’ll also load the R/broman package for some utilities and plotting functions, and R/qtlcharts for interactive graphs.

We will follow the steps by `Karl Broman` found [here](https://kbroman.org/qtl2/assets/vignettes/do_diagnostics.html)

### Loading Project
```{r generate json file for all batches, eval=TRUE}

gm <- read_cross2(paste0(filepaths,"/qtl2/gm.json"))
#gmf <- read_cross2("/Users/corneb/Documents/MyJax/CS/Projects/Oconnor/genoprobs/3.batches/all.samples/qtl2/gm_sorted.json")

#gm <- get(load("gm_DO_3batches.RData"))
#gmf

##pruning out controls

#control.mice <- c("KJgDNA2", "KJgDNA3", "KJgDNA4", "KJgDNA5", "KJgDNA6", "KJgDNA7")

#gm <- gmf[! ind_ids(gmf) %in%control.mice,]
#gm <- gmf[! sapply(strsplit(as.character(ind_ids(gmf)),'_',fixed=T),function(x) x[6]) %in%control.mice,]
gm

sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
#sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
#samples.all <- read.csv(sample_file)
#samples <- samples.all[! samples.all$Original.Mouse.ID %in%control.mice,]
samples <- read.csv(sample_file)
all.equal(as.character(ind_ids(gm)), as.character(samples$Unique.Sample.ID))

````

### Missing Data
```{r Missing data per sample, fig.height = 6, fig.width = 9.5, fig.align = "center"}

percent_missing <- n_missing(gm, "ind", "prop")*100

labels <- paste0(as.character(do.call(rbind.data.frame, strsplit(names(percent_missing), "_"))[,7]), " (", round(percent_missing,2), "%)")
iplot(seq_along(percent_missing), percent_missing, indID=labels,
      chartOpts=list(xlab="Mouse", ylab="Percent missing genotype data",
                     ylim=c(0, 70)))

#save into pdf
pdf(file = "output/Percent_missing_genotype_data_3.batches.pdf", width = 20, height = 20)
#labels <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "V01_"))[,2])
labels <- as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])
#labels <- paste0(names(percent_missing), " (", round(percent_missing,2), "%)")
labels[percent_missing < 10] = ""
# Change point shapes and colors
p <- ggplot(data = data.frame(Mouse=seq_along(percent_missing),  
                         Percent_missing_genotype_data = percent_missing,
                         batch = factor(as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,6]))
                         #batch = factor(as.character(do.call(rbind.data.frame, strsplit(as.character(samples$Directory), "_"))[,5]))
                         ), 
        aes(x=Mouse, y=Percent_missing_genotype_data, color = batch)) +
        geom_point() +
        geom_hline(yintercept=10, linetype="solid", color = "red") +
        geom_text_repel(aes(label=labels), vjust = 0, nudge_y = 0.01, show.legend = FALSE, size=3) +
        theme(text = element_text(size = 10))
p

dev.off()

p

save(percent_missing,file = "data/percent_missing_id_3.batches.RData")

gm.covar = data.frame(id=rownames(gm$covar),gm$covar)
qc_info_cr <- merge(gm.covar,
                  data.frame(id = names(percent_missing),percent_missing = percent_missing,stringsAsFactors = F),by = "id")
bad.sample.cr <- qc_info_cr[qc_info_cr$percent_missing >= 10,]

```


```{r high missing table, echo=FALSE}

rownames(bad.sample.cr) <- NULL
bad.sample.cr[] <- lapply(bad.sample.cr, as.character)
bad.sample.cr$Neogen_Sample_ID <- sapply(strsplit(as.character(bad.sample.cr$id),'_',fixed=T),function(x) x[6])
bad.sample.cr$Genoprobs_Sample_ID <- bad.sample.cr$id

bad.sample.cr[c("Neogen_Sample_ID","Genoprobs_Sample_ID","percent_missing")] %>%
  kable(escape = F,align = c("ccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) #%>%
  #column_spec(1:3, width = "5cm") 

```

```{r removing bad samples CR, include=FALSE}

##removing bad samples
#gm <- gm[paste0("-",as.character(bad.sample.cr$id)),]#

#gm

```

### Sex
```{r Sexes, fig.height = 6, fig.width = 9.5, fig.align = "center"}
hdf5_filename <- dir(path = filepaths, pattern = "^hdf5_*", full.names = TRUE)
snps_file <- "/Users/corneb/Documents/MyJax/CS/Projects/QTL2_support.files/gm_uwisc_v1.csv"
snps <- read.csv(snps_file)

snps <- snps[snps$unique == TRUE, ]
#snps <- snps[snps$chr %in% c(1:19, "X"), ]
snps$chr <- sub("^chr", "", snps$chr)  ###remove prefix "chr"
colnames(snps)[colnames(snps)=="bp_mm10"] <- "pos" 
colnames(snps)[colnames(snps)=="cM_cox"] <- "cM"
snps <- snps %>% drop_na(chr, marker) 
snps$pos <- snps$pos * 1e-6
rownames(snps) <- snps$marker
colnames(snps)[1:4] <- c("marker", "chr", "pos", "pos") 

#  g <- h5read(hdf5_filename, "G")
#  g <- do.call(cbind, g)
x <- h5read(hdf5_filename, "X") # X channel intensities
x <- do.call(cbind, x)
y <- h5read(hdf5_filename, "Y") # Y channel intensities
y <- do.call(cbind, y)
rn <- h5read(hdf5_filename, "rownames")[[1]]  # markers 
cn <- h5read(hdf5_filename, "colnames")  # samples
cn <- do.call(c, cn)
# dimnames(g) <- list(rn, cn)
dimnames(x) <- list(rn, cn)
dimnames(y) <- list(rn, cn)
#cr <- colMeans(g != "--") # Call rate for each sample avg 0.95
#  sex <- determine_sex(x = x, y = y, markers = snps)$se

markers <- snps

chrx <- markers$marker[which(markers$chr == "X")]
chry <- markers$marker[which(markers$chr == "Y")]
#x[chrx,ind_ids(gm)]

chrx_int <- colMeans(x[chrx,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])] + y[chrx,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])], na.rm = T)
chry_int <- colMeans(x[chry,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])] + y[chry,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])], na.rm = T)
#  df <- data.frame(x = chrx_int, y = chry_int)
#  mc <- Mclust(data = df, G = 2, model = "VVV", prior=priorControl()) 
#  # Determine which cluster is females by selecting the cluster with the
#  # higher Chr X intensity mean.
#  cluster_means <- mc$parameters$mean
#  female_cl <- which.max(cluster_means["x",])
#  sex <- rep("M", length(mc$classification))
#  names(sex) <- colnames(x)
#  sex[mc$classification == female_cl] <- "F"
#  return(list(sex = sex))

#xint <- read_csv_numer("/Users/corneb/Documents/MyJax/CS/Projects/Simon.John.Tolman/Data/GM/Simon.John.Tolman_qtl2_chrXint.csv", transpose=TRUE)
#yint <- read_csv_numer("data/Univ_of_Penn_Thaiss_qtl2_chrYint.csv", transpose=TRUE)

#xint2 <- t(x[chrx,ind_ids(gm)])
#yint2 <- t(y[chry,ind_ids(gm)])

#sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
#samples <- read.csv(sample_file)

all.equal(as.character(ind_ids(gm)), as.character(samples$Unique.Sample.ID))

#sex order
samples$Sex <- 'F'
sex <- samples$Sex
#sex <- gm$covar[rownames(xint),"sexp"]
#sex <- substr(rownames(xint), 1, 1)

#x_pval <- apply(xint2, 2, function(a) t.test(a ~ sex)$p.value)
#y_pval <- apply(yint2, 2, function(a) t.test(a ~ sex)$p.value)

#xint_ave <- rowMeans(xint2[, x_pval < 0.05/length(x_pval)], na.rm=TRUE)
#yint_ave <- rowMeans(yint2[, y_pval < 0.05/length(y_pval)], na.rm=TRUE)

#BAF = yint / (xint + yint)
#LRR = log2((xint + yint) / (xint_ave+yint_ave))

point_colors <- as.character( brocolors("web")[c("green", "purple")] )
percent_missing <- n_missing(gm, summary="proportion")*100
labels <- paste0(names(chrx_int), " (", round(percent_missing), "%)")
iplot( chrx_int,  chry_int, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Average Y chr intensity"))
```

For figures above and below, those labelled as female in metadata given, are coloured `green`, with those labelled as male are coloured as `purple`. The above is an interactive scatterplot of the average SNP intensity on the Y chromosome versus the average SNP intensity on the X chromosome.

The cluster in the upper-left are male mice (low X chromosome intensity and high Y chromosome intensity), and the cluster in the lower-right are female mice (high X chromosome intensity and low Y chromsome intensity).  

```{r Sexes p2, fig.height = 6, fig.width = 9.5, fig.align = "center"}

phetX <- rowSums(gm$geno$X == 2)/rowSums(gm$geno$X != 0)
phetX <- phetX[as.character(do.call(rbind.data.frame, strsplit(names(phetX), "_"))[,7]) %in% names(chrx_int)]
names(phetX) <- as.character(do.call(rbind.data.frame, strsplit(names(phetX), "_"))[,7])
iplot(chrx_int, phetX, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Proportion het on X chr"))
```

In the above scatterplot, we show the proportion of hets vs the average intensity for the X chromosome SNPs. In calculating the proportion of heterozygous genotypes for the individuals, we look at X chromosome genotypes equal to 2 which corresponds to the heterozygote) relative to not being 0 (which is used to encode missing genotypes). The genotypes are arranged with rows being individuals and columns being markers. 

The following are the mice that have had sex incorrectly assigned:

```{r sex diff table, echo=FALSE}

bad.sample.sex <- subset(samples, (samples$Inferred.Sex != substring(samples$Sex, 1, 1)) | is.na(substring(samples$Sex, 1, 1)))
#bad.sample.sex$Sex[is.na(bad.sample.sex$Sex)] <- '--'
bad.sample.sex$Neogen_Sample_ID <- sapply(strsplit(as.character(bad.sample.sex$Unique.Sample.ID),'_',fixed=T),function(x) x[7])
bad.sample.sex$Genoprobs_Sample_ID <- bad.sample.sex$Unique.Sample.ID
rownames(bad.sample.sex) <- NULL

bad.sample.sex[c("Neogen_Sample_ID","Genoprobs_Sample_ID","Sex","Inferred.Sex")] %>%
  kable(escape = F,align = c("ccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(1:4, width = "5cm") 

```

### Sample Duplicates
```{r Sample duplicates}

cg <- compare_geno(gm, cores=10)
summary.cg <- summary(cg)
```
```{r Sample duplicates table, echo=FALSE}
rownames(summary.cg) <- NULL
summary.cg$iid1 <- sapply(strsplit(as.character(summary.cg$ind1),'_',fixed=T),function(x) x[7]) 
summary.cg$iid2 <- sapply(strsplit(as.character(summary.cg$ind2),'_',fixed=T),function(x) x[7]) 
summary.cg$pc.miss.ind1 <- percent_missing[match(summary.cg$ind1, names(percent_missing))]
summary.cg$pc.miss.ind2 <- percent_missing[match(summary.cg$ind2, names(percent_missing))]

#summary.cg$iid1 <- summary.cg$ind1
#summary.cg$iid2 <- summary.cg$ind2

summary.cg[c(9,1,11,10,2,12,3:6)] %>%
  kable(escape = F,align = c("llcccccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(1:8, width = "5cm") 

```

Here is a histogram of the proportion of matching genotypes. The tick marks below the histogram indicate individual pairs.

```{r Sample duplicates figures}
save(summary.cg,file = "data/summary.cg_3.batches.RData")

#pdf(file = "output/Proportion_matching_genotypes_before_removal_of_bad_samples_3.batches.pdf", width = 20, height = 20) 
#par(mar=c(5.1,0.6,0.6, 0.6))
#hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
#     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
#rug(cg[upper.tri(cg)])
#dev.off()

par(mar=c(5.1,0.6,0.6, 0.6))
hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cg[upper.tri(cg)])

```

```{r Sample duplicates 2 figures, include=FALSE}

cgsub <- cg[percent_missing < 50, percent_missing < 50]
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cgsub[upper.tri(cgsub)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes [percent missing < 50%]")
rug(cgsub[upper.tri(cgsub)])

```

```{r top 20 high missing rate, include = FALSE}
##show top 20 samples with missing genotypes
#percent_missing <- n_missing(gm, "ind", "prop")*100
#round(sort(percent_missing, decreasing=TRUE)[1:20], 1)  
```

```{r removing duplicate samples, include=FALSE}
#removing duplcate pairs

#pheno_file_nick <- dir(path = filepathp, pattern = "*_pheno.csv+$", full.names = TRUE)
#pheno.pi <- read.csv(pheno_file_nick)

#master <- as.data.frame(read_excel("/Users/corneb/Documents/MyJax/CS/Projects/Simon.John.Tolman/Data/meta.data/inventory_file.xlsx", sheet = "inventory_file"))

#sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
#samples <- read.csv(sample_file)

#project <- read.csv("/Users/corneb/Documents/MyJax/CS/Projects/Simon.John.Tolman/haplo.reconst/2020-11-09/all.samples/all.samples/project.csv", as.is=T)
#project$dir2 <- sapply(strsplit(project$dir,'/',fixed=T),function(x) x[7])

#inventory <- master[master$"Sample ID...4" %in% samples$Original.Mouse.ID,]
#invent.dir2 <- merge(inventory,project, by.x=c("Directory"), by.y=c("dir2"), all=T)
#invent.dodb <- merge(invent.dir2,samples, by.x=c("Directory","Sample ID...4"), by.y=c("Directory","Original.Mouse.ID"),all=T)

#gm.pheno1 <- merge(invent.dodb,pheno.pi, by.x=c("Sample ID...5"), by.y=c("Mouse.ID"), all.x=T)
#gm.pheno <- gm.pheno1[c("Correct.Mouse.ID","Unique.Sample.ID","Unique Sample ID","Sample ID...4","Sample ID...5","Directory",
#                     "Sex.x","Sex.y","Sex","Inferred.Sex","Birth Date","DOB","DO Generation","DO.Mother.Generation","DO.Generation","Strain.x","Strain.y",
#                     "Lmx1b allele","Lmx1b.allele","Lmx1b Genotype","Lmx1b.Genotype","Parental.Strain.GigaMUGA.ID","Parental Strain GigaMUGA ID",
#                     "Run.Date","Geneseek.Sample.ID","Include","cr","Max_IOP")]

#summary.cg$Name.ind1 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind1), "_"))[,6])
#summary.cg$Name.ind2 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind2), "_"))[,6])
summary.cg$Name.ind1 <- summary.cg$ind1
summary.cg$Name.ind2 <- summary.cg$ind2
summary.cg$miss.ind1 <- percent_missing[match(summary.cg$ind1, names(percent_missing))]
summary.cg$miss.ind2 <- percent_missing[match(summary.cg$ind2, names(percent_missing))]
#summary.cg$phen.ind1 <- gm.pheno$Max_IOP[match(summary.cg$ind1, gm.pheno$Unique.Sample.ID)]
#summary.cg$phen.ind2 <- gm.pheno$Max_IOP[match(summary.cg$ind2, gm.pheno$Unique.Sample.ID)]
summary.cg$remove.id <- ifelse((summary.cg$miss.ind1 > summary.cg$miss.ind2), summary.cg$ind1, summary.cg$ind2)
#summary.cg$remove.id.p1 <- ifelse(is.na(summary.cg$phen.ind1) & (is.na(summary.cg$phen.ind1) | !is.na(summary.cg$phen.ind1)), summary.cg$ind1, '')
#summary.cg$remove.id.p2 <- ifelse(is.na(summary.cg$phen.ind2) & (is.na(summary.cg$phen.ind1) | !is.na(summary.cg$phen.ind1)), summary.cg$ind2, '')
#remove.IDs <- rbind(summary.cg$remove.id.m[-1],summary.cg$remove.id.p1[-1],summary.cg$remove.id.p2[-1])
summary.cg$remove.id  

#qc_info$remove.id.duplicated <- ifelse(qc_info$id %in% summary.cg$remove.id, TRUE,FALSE)

#bad.sample.cr.dup <- NULL

#bad.sample.cr.dup$id <- c("Univ_of_Penn_Thaiss_MURMUGV01_20200130_9_A6",#
#                      "Univ_of_Penn_Thiass_MURGIGV01_20200226_90_C12", 
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_109_G2",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_111_B3",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_110_H5",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_112_F5")

#gm <- gm[paste0("-",as.character(bad.sample.cr.dup$id)),]

#gm

#percent_missing <- n_missing(gm, "ind", "prop")*100
#round(sort(percent_missing, decreasing=TRUE)[1:19], 1)

```

### Array Intensities
```{r Array intensities, fig.height = 6, fig.width = 9.5, fig.align = "center"}

#load the intensities.fst_3.batches.RData
#load("data/intensities.fst_3.batches.RData")

xn <- x[,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])]
xn <- xn[snps$marker,]
xnm <- rownames(xn)

yn <- y[,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])]
yn <- yn[snps$marker,]

# bring together in one matrix
result <- cbind(snp=rep(snps$marker, 2),
                channel=rep(c("x", "y"), each=length(snps$marker)),
                as.data.frame(rbind(xn, yn)))
rownames(result) <- 1:nrow(result)

# bring SNP rows together
result <- result[as.numeric(t(cbind(seq_along(snps$marker), seq_along(snps$marker)+length(snps$marker)))),]
rownames(result) <- 1:nrow(result)

#load the intensities.fst_3.batches.RData
#load("data/heh/intensities.fst_3.batches.RData")
#X and Y channel
X <- result[result$channel == "x",]
rownames(X) <- X$snp
X <- X[,c(-1,-2)]

Y <- result[result$channel == "y",]
rownames(Y) <- Y$snp
Y <- Y[,c(-1,-2)]

int <- result

#int <- result

#rm(result)
int <- int[seq(1, nrow(int), by=2),-(1:2)] + int[-seq(1, nrow(int), by=2),-(1:2)]
int <- int[,intersect(as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7]), colnames(int))]
names(percent_missing) <- as.character(do.call(rbind.data.frame, strsplit(names(percent_missing), "_"))[,7])
n <- names(sort(percent_missing[intersect(as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7]), colnames(int))], decreasing=TRUE))
iboxplot(log10(t(int[,n])+1), orderByMedian=FALSE, chartOpts=list(ylab="log10(SNP intensity + 1)"))

```
In the above plot, distributions of array intensities (after a log10(x+1) transformation) are displayed. 

The arrays are sorted by the proportion of missing genotype data for the sample, and the curves connect various quantiles of the intensities.

```{r Array intensities percentile figure, fig.height = 6, fig.width = 9.5, fig.align = "center"}

qu <- apply(int, 2, quantile, c(0.01, 0.99), na.rm=TRUE)
group <- (percent_missing >= 19.97) + (percent_missing > 5) + (percent_missing > 2) + 1
labels <- paste0(colnames(qu), " (", round(percent_missing), "%)")
iplot(qu[1,], qu[2,], indID=labels, group=group,
      chartOpts=list(xlab="1 %ile of array intensities",
                     ylab="99 %ile of array intensities",
                     pointcolor=c("#ccc", "slateblue", "Orchid", "#ff851b")))
```
For this particular set of arrays, a plot of the 1 %ile vs the 99 %ile is quite revealing. In the following, the orange points are those with > 20% missing genotypes, the pink points are the samples with 5-20% missing genotypes, and the blue points are the samples with 2-5% missing genotypes.

### Genotype Frequencies

The following triangle plots show the genotype frequency distributions for the mice, among the four groups of markers with common minor allele frequency (MAF) in the founder strains. These plots make use of the fact that for a point within an equilateral triangle, the sum of the distances to the three sides is a constant.

```{r Genotype frequencies}
g <- do.call("cbind", gm$geno[1:19])
fg <- do.call("cbind", gm$founder_geno[1:19])
g <- g[,colSums(fg==0)==0]
fg <- fg[,colSums(fg==0)==0]
fgn <- colSums(fg==3)

gf_ind <- vector("list", 4)
for(i in 1:4) {
  gf_ind[[i]] <- t(apply(g[,fgn==i], 1, function(a) table(factor(a, 1:3))/sum(a != 0)))
}

par(mfrow=c(4,1), mar=c(0.6, 0.6, 2.6, 0.6))
for(i in 1:4) {
  triplot(c("AA", "AB", "BB"), main=paste0("MAF = ", i, "/8"))
  tripoints(gf_ind[[i]], pch=21, bg="lightblue")
  tripoints(c((1-i/8)^2, 2*i/8*(1-i/8), (i/8)^2), pch=21, bg="violetred")
  
  if(i>=3) { # label mouse with lowest het
    wh <- which(gf_ind[[i]][,2] == min(gf_ind[[i]][,2]))
    tritext(gf_ind[[i]][wh,,drop=FALSE] + c(0.02, -0.02, 0),
            names(wh), adj=c(0, 1))
  }
  
  # label other mice
  if(i==1) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.3]
  }
  else if(i==2) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.48]
  }
  else if(i==3) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.51]
  }
  else if(i==4) {
    lab <- rownames(gf_ind[[i]])[gf_ind[[i]][,2]>0.6]
  }
  
  for(ind in lab) {
    if(grepl("^F", ind) && i != 3) {
      tritext(gf_ind[[i]][ind,,drop=FALSE] + c(-0.01, 0, +0.01), ind, adj=c(1,0.5))
    } else {
      tritext(gf_ind[[i]][ind,,drop=FALSE] + c(0.01, 0, -0.01), ind, adj=c(0,0.5))
    }
  }
}

````

The majority of individuals are tightly clustered around the expected distribution (in pink). There are a few outliers with elevated & reduced heterozygosity, but most of these are among the samples with highest rates of missing genotypes. 

```{r removing low af, include=FALSE}

#bad.sample.cr.dup.gf <- NULL

#bad.sample.cr.dup.gf$id <- c("Univ_of_Penn_Thaiss_MURMUGV01_20200130_108_D12")

#gm <- gm[paste0("-",as.character(bad.sample.cr.dup.gf$id)),]

#gm

```

### Crossover Counts
#```{r Crossover counts, fig.height = 6, fig.width = 9.5, fig.align = "center"}
#load pre-caluated results
#load("data/pr_3.batches.RData")
#pr<-readRDS(paste0(filepaths,"cell_gig_data__GigaMUGA_genoprobs_36state.rds"))
#load("data/m_3.batches.RData")
#load("data/nxo_3.batches.RData")

#m <- maxmarg(pr)
#nxo <- count_xo(m)
#save(pr, file = "data/pr_3.batches.RData")
#str(pr)
#save(m, file = "data/m_3.batches.RData")
#save(nxo, file = "data/nxo_3.batches.RData")

#e <- calc_errorlod(gm, pr)
#str(e)
#e <- do.call("cbind", e)
#save(e, file = "data/e_3.batches.RData")

#snpg <- predict_snpgeno(gm, m)
#snpg <- do.call("cbind", snpg)
#save(snpg, file = "data/snpg_3.batches.RData")

#crossover
#totxo <- rowSums(nxo)[rownames(gm$covar)]
#names(totxo) <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "_"))[,6])
#iplot(seq_along(totxo),
#      totxo,
#      group=gm$covar$generation,
#      chartOpts=list(xlab="Mouse", ylab="Number of crossovers", 
#                     margin=list(left=80,top=40,right=40,bottom=40,inner=5),
#                     axispos=list(xtitle=25,ytitle=50,xlabel=5,ylabel=5)))

#save crossover into pdf
#pdf(file = "output/number_crossover_3.batches.pdf")
#cross_over <- data.frame(Mouse = seq_along(totxo), id=names(totxo), Number_crossovers = totxo, generation = gm$covar$generation)
#names(totxo) <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "V01_"))[,2])
#names(totxo) <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "_"))[,6])
#names(totxo)[totxo >= 200 & totxo <= 800] = ""
# Change point shapes and colors
#p <-ggplot(cross_over, aes(x=Mouse, y=Number_crossovers, fill = generation, color=generation)) +
#  geom_point() +
#  geom_text_repel(aes(label=names(totxo),hjust=0,vjust=0), show.legend = FALSE)
#p
#dev.off()

#p

```
Counts of inferred crossovers can be a useful diagnostic for identifying problem samples, which may show an excessive number of apparent crossovers. Above is a plot of the number of crossovers vs the mouse index, colored by generation.

We’re looking for mice with an excessive number of crossovers as well as mice with very few crossovers.  These are shown below

**NB:** DO generation of an individual is taken into account. 

#```{r  corssover counts, echo = FALSE}
#Here are the crossover counts for those  mice:
#tmp <- cbind(percent_missing=round(percent_missing[as.character(do.call(rbind.data.frame, strsplit(rownames(gm$covar), "_"))[,7])],2), total_xo=totxo)[percent_missing[as.character(do.call(rbind.data.frame, strsplit(rownames(gm$covar), "_"))[,7])] >= 5,]
#tmp[order(tmp[,1]),]

##number of crossovers

#totxo <- rowSums(nxo)[rownames(gm$covar)]
#totxot <- totxo[totxo <= 200 | totxo >= 1000]

#totxot <- as.data.frame(totxot)
##totxot <- as.data.frame(t(totxot))

#totxot$Neogen_Sample_ID <- sapply(strsplit(rownames(totxot),'_',fixed=T),function(x) x[6]) 
#totxot$Genoprobs_Sample_ID <- rownames(totxot)
#totxot$number.xo <- totxot$totxot
#totxot$id <- rownames(totxot)

#rownames(totxot) <- NULL

#totxot[c("Neogen_Sample_ID","Genoprobs_Sample_ID","number.xo")] %>%
#  kable(escape = F,align = c("cc"),linesep ="\\hline") %>%
#  kable_styling(full_width = F) %>%
#  kable_styling("striped", full_width = F) 

```

#### Genotyping Error LOD Scores
#```{r  Genotyping error LOD scores, fig.height = 6, fig.width = 9.5, fig.align = "center"}
#load("data/e_3.batches.RData")
#errors_ind <- rowSums(e>2)[rownames(gm$covar)]/n_typed(gm)*100
#lab <- paste0(as.character(do.call(rbind.data.frame, strsplit(names(errors_ind), "_"))[,7]), " (", myround(percent_missing[as.character(do.call(rbind.data.frame, strsplit(rownames(gm$covar), "_"))[,7])],1), "%)")
#iplot(seq_along(errors_ind), errors_ind, indID=lab,
#      chartOpts=list(xlab="Mouse", ylab="Percent genotyping errors", ylim=c(0, 15),
#                     axispos=list(xtitle=25, ytitle=50, xlabel=5, ylabel=5)))
#save(errors_ind, file = "data/errors_ind_3.batches.RData")

```

### Apparent Genotyping Errors
#```{r apparent genotyping errors, fig.height = 6, fig.width = 9.5, fig.align = "center"}

#load("data/snpg_3.batches.RData")
#snpg <- gm$geno

#snpgn <- snpg[rownames(gm$covar),]
#snpgn <- do.call("cbind", gm$geno)

#gobs <- do.call("cbind", gm$geno)
#gobs[gobs==0] <- NA

#par(pty="s")
#err_direct <- rowMeans(snpgn != gobs, na.rm=TRUE)*100
#errors_ind_0 <- rowSums(e > 0)[rownames(gm$covar)]/n_typed(gm)*100
#par(mar=c(4.1,4.1,0.6, 0.6))
#grayplot(errors_ind_0, err_direct,
#         xlab="Percent errors (error LOD > 0)",
#         ylab="Percent errors (obs vs predicted)",
#         xlim=c(0, 35), ylim=c(0, 35))
#abline(0,1,lty=2, col="gray60")

#pdf(file = "output/Percent_genotype_errors_obs_vs_predicted_3.batches.pdf",width = 20, height = 20) 
#par(pty="s")
#err_direct <- rowMeans(snpgn != gobs, na.rm=TRUE)*100
#errors_ind_0 <- rowSums(e > 0)[rownames(gm$covar)]/n_typed(gm)*100
#par(mar=c(4.1,4.1,0.6, 0.6))
#grayplot(errors_ind_0, err_direct,
#         xlab="Percent errors (error LOD > 0)",
#         ylab="Percent errors (obs vs predicted)",
#         xlim=c(0, 35), ylim=c(0, 35))
#abline(0,1,lty=2, col="gray60")
#dev.off()

#par(pty="s")
#par(mar=c(4.1,4.1,0.6, 0.6))
#grayplot(errors_ind_0, err_direct,
#         xlab="Percent errors (error LOD > 0)",
#         ylab="Percent errors (obs vs predicted)",
#         xlim=c(0, 35), ylim=c(0, 35))
#abline(0,1,lty=2, col="gray60")

```


### Removing Samples
```{r removing samples }

##percent missing
gm.covar = data.frame(id=as.character(do.call(rbind.data.frame, strsplit(rownames(gm$covar), "_"))[,7]),gm$covar)
qc_info <- merge(gm.covar,
                  data.frame(id = names(percent_missing),percent_missing = percent_missing,stringsAsFactors = F),by = "id")

#cross_over
#qc_info <- merge(qc_info,
#                 data.frame(id = rownames(cross_over),
#                            Number_crossovers = cross_over$Number_crossovers,stringsAsFactors = F),by = "id")

#missing sex
#qc_info$sex.match <- ifelse(qc_info$sexp == qc_info$sex, TRUE, FALSE)
rownames(samples) <- as.character(samples$Original.Mouse.ID)
samples <- samples[as.character(qc_info$id),]
#samples$Unique.Sample.ID <- as.character(samples$Unique.Sample.ID)
all.equal(as.character(qc_info$id), as.character(samples$Original.Mouse.ID))
qc_info$sex.match <- ifelse(samples$Inferred.Sex == substring(samples$Sex, 1, 1), TRUE, FALSE)


#genotype errors
#qc_info <- merge(qc_info,
#                 data.frame(id = as.character(do.call(rbind.data.frame, strsplit(names(errors_ind), "_"))[,7]),
#                            genotype_erros = errors_ind,stringsAsFactors = F),by = "id")

##duplicated id to be remove
qc_info$duplicate.id <- ifelse(qc_info$id %in% as.character(do.call(rbind.data.frame, strsplit(summary.cg$remove.id, "_"))[,7]), TRUE,FALSE)

#bad.sample <- qc_info[qc_info$generation ==1 | qc_info$Number_crossovers <= 200 | qc_info$Number_crossovers >=1000 | qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 1 | qc_info$remove.id.duplicated == TRUE,]
bad.sample <- qc_info[qc_info$percent_missing >= 10 | qc_info$duplicate.id == TRUE,]

save(qc_info, bad.sample, file = "data/qc_info_bad_sample_3.batches.RData")

gm_samqc <- gm[paste0("-",as.character(bad.sample$id.1)),]

#fix sex mismatch
#gm_samqc$covar[gm_samqc$covar$id %in% names(yint_ave[!names(yint_ave) %in% bad.sample$id][yint_ave[!names(yint_ave) %in% bad.sample$id] <= 0.1]),"sexp"] <- "F"

gm_samqc

save(gm_samqc, file = "data/gm_samqc_3.batches.RData")

# update other stuff
#e <- e[ind_ids(gm_samqc),]
#g <- g[ind_ids(gm_samqc),]
#snpg <- snpg[ind_ids(gm_samqc),]

#save(e,g,snpg, file = "data/e_g_snpg_samqc_3.batches.RData")
#save(e,g, file = "data/e_g_snpg_samqc_3.batches.RData")

```

#Here is the list of samples that were removed:

#```{r removing table, echo=FALSE}

#removed.ids <- bad.sample[c(2)]
#names(removed.ids) <- "Genoprobs_Sample_ID"
#removed.ids$Neogen_Sample_ID <-  as.character(do.call(rbind.data.frame, strsplit(as.character(removed.ids$Genoprobs_Sample_ID), "_"))[,7])

#rownames(removed.ids) <- NULL

#removed.ids %>%
#  kable(escape = F,align = c("c"),linesep ="\\hline") %>%
#  kable_styling(full_width = F) %>%
#  kable_styling("striped", full_width = F) 

#```

Below is a table summarising the problematic samples found throughout QC. These include the following:

+ no_pheno == sample does not have a phenotype (if applicable)
+ high_miss == sample has higher than 10% missing genotypes 
+ diff_sex == sample has sex mismatch
+ high_xo == sample had crossover events greater than 1000 and less than 200
+ high_geno_errors == Sample has geno errors above 1%
+ DO_generation == DO generation is 1 (if applicable)
+ dublicate_ID == sample was duplicated

**NB:** For duplcate pairs, the one that was chosen to be removed was the one that had a higher missing rate

```{r removing samples table, echo=FALSE}


#pheno_file_nick <- dir(path = filepathp, pattern = "*_pheno.csv+$", full.names = TRUE)
#gm.pheno <- read.csv(pheno_file_nick)
#gm.pheno$id <-  gm.pheno$Unique.Sample.ID
#qc_info <- merge(qc_info,gm.pheno,by = "id",all.x=T)

#bad.sample <- qc_info[qc_info$ngen ==1 | qc_info$Number_crossovers <= 200 | qc_info$Number_crossovers >=1000 | qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 1 | qc_info$remove.id.duplicated == TRUE | is.na(qc_info$Time_min)==TRUE | is.na(qc_info$Distance_m)==TRUE | is.na(qc_info$Energy_J)==TRUE,]

ind1<-summary.cg[c("ind1","iid1")]
names(ind1) <- c("iid","id")
ind2<-summary.cg[c("ind2","iid2")]
names(ind2) <- c("iid","id")
dup.IDS <- rbind(ind1, ind2)
dup.IDS <-dup.IDS[!duplicated(dup.IDS),]
dup.IDS$duplicated <- "TRUE"

bad.sample <- merge(qc_info,dup.IDS,by=c("id"),all=T)
bad.sample[is.na(bad.sample)] <- ""

bad <- NULL
#bad$Mouse_ID <- bad.sample$"Sample ID...5"
#bad<-as.data.frame(bad)
#bad$Neogen_Sample_ID <- bad.sample$"Sample ID...4"
bad$Genoprobs_Sample_ID <- as.character(bad.sample$id.1)
bad$Neogen_Sample_ID <- as.character(do.call(rbind.data.frame, strsplit(as.character(bad$Genoprobs_Sample_ID), "_"))[,7])
#bad$Unique_Sample_ID <- bad.sample$id
#bad$no_pheno <- ifelse(is.na(bad.sample$Max_IOP)==TRUE, 'XX', '')
bad$high_miss <- ifelse(bad.sample$percent_missing >= 10, 'XX', '')
bad$diff_sex <- ifelse(bad.sample$sex.match ==FALSE | is.na(bad.sample$sex.match) == TRUE, 'XX', '')
bad$high_xo <- ifelse(bad.sample$Number_crossovers <= 200 | bad.sample$Number_crossovers >=1000, 'XX', '')
bad$DO.Generation <- ifelse(bad.sample$generation ==1, 'XX', '')
bad$high_geno.errors <- ifelse(bad.sample$genotype_erros >= 1, 'XX', '')
bad$duplicate_ID <- ifelse(bad.sample$duplicated == TRUE, 'XX', '')
bad[is.na(bad)] <- ""
bad[bad=='NA'] <- ""
df <- do.call(cbind, bad)
bad <- as.data.frame(df)

badind <- subset(bad, 
         #bad$no_pheno == 'XX'|
         bad$high_miss == 'XX'|
         bad$diff_sex == 'XX'|  
#         bad$high_xo == 'XX'|
#         bad$high_geno.errors == 'XX'|
#         bad$DO.Generation == 'XX'|
         bad$duplicate_ID == 'XX')


#badind <- bad[bad$no_pheno == 'XX',]

badind[] <- lapply(badind, as.character)
#badind$Thaiss_ID <- ifelse(badind$Thaiss == 994 | badind$Thaiss == 995 | badind$Thaiss == 996 |badind$Thaiss == 997 | badind$Thaiss == 998 | badind$Thaiss == 999, "--", bad$Thaiss_ID)

rownames(badind) <- NULL

badind[c(1:4,6)] %>% 
   mutate(
     #no_pheno = ifelse(no_pheno == 'XX',
     #             cell_spec(no_pheno, color = 'green',background = 'green'),
     #             ''),
     high_miss = ifelse(high_miss == 'XX',
                  cell_spec(high_miss, color = 'red',background = 'red'),
                  ''),
     diff_sex = ifelse(diff_sex == 'XX',
                  cell_spec(diff_sex, color = 'blue',background = 'blue'),
                  ''),
#     high_xo = ifelse(high_xo == 'XX',
#                  cell_spec(high_xo, color = 'purple',background = 'purple'),
#                  ''),
#     high_geno.errors = ifelse(high_geno.errors == 'XX',
#                  cell_spec(high_geno.errors, color = 'pink',background = 'pink'),
#                  ''),
#     DO.Generation = ifelse(DO.Generation == 'XX',
#                  cell_spec(DO.Generation, color = 'lightgreen',background = 'lightgreen'),
#                  ''),
     duplicate_ID = ifelse(duplicate_ID == 'XX',
                  cell_spec(duplicate_ID, color = 'orange',background = 'orange'),
                  '')
     ) %>%
   kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
   kable_styling("striped", full_width = F) %>%
   #column_spec(2:6, width = "3cm") %>%
   column_spec(1, width = "5cm")

```

