---
title: "Step 1: Sample QC [7.batches_myo]"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

filepathp = '/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output_7.batches_myo/'
filepaths = '/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output_7.batches_myo/'
#setwd(filepaths)
setwd("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/qc/workflowr/Serreze-T1D_Workflow")

#libraries
library(broman)
library(qtl2)
library(qtlcharts)
library(ggplot2)
library(ggrepel)
#library(DOQTL)
library(mclust)
#source("code/reconst_utils.R")
library("kableExtra")
library("knitr")
library("fst")
library(data.table) 
library(tidyr)     
library(mclust)     
library(rhdf5)      
library(optparse)
library(dplyr)
library(cluster)
library(rhdf5)
library(ggplot2)
library(readxl) 
library(lubridate)

```
This script is running genotype QC on raw data (with some outcomes already seen in the [project at a glace](summary.html)). Here, we first load the R/qtl2 package and the data.
We’ll also load the R/broman package for some utilities and plotting functions, and R/qtlcharts for interactive graphs.

We will follow the steps by `Karl Broman` found [here](https://kbroman.org/qtl2/assets/vignettes/do_diagnostics.html)

### Loading Project
```{r generate json file for all batches, eval=TRUE}


gm <- get(load("data/gm_serreze_bc_7.batches_myo.RData"))
gm

sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
samples <- read.csv(sample_file)
all.equal(as.character(ind_ids(gm)), as.character(samples$Original.Mouse.ID))
all.equal(as.character(ind_ids(gm)), as.character(samples$Unique.Sample.ID))

````

### Missing Data
```{r Missing data per sample, fig.height = 6, fig.width = 9.5, fig.align = "center"}

percent_missing <- n_missing(gm, "ind", "prop")*100

labels <- paste0(as.character(do.call(rbind.data.frame, strsplit(names(percent_missing), "_"))[,7]), " (", round(percent_missing,2), "%)")
#labels <- paste0(names(percent_missing), " (", round(percent_missing,2), "%)")
iplot(seq_along(percent_missing), percent_missing, indID=labels,
      chartOpts=list(xlab="Mouse", ylab="Percent missing genotype data",
                     ylim=c(0, 70)))

#save into pdf
pdf(file = "output/Percent_missing_genotype_data_7.batches_myo.pdf", width = 20, height = 20)
#labels <- as.character(do.call(rbind.data.frame, strsplit(names(totxo), "V01_"))[,2])
labels <- as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])
#labels <- paste0(names(percent_missing), " (", round(percent_missing,2), "%)")
#labels <- ind_ids(gm)
labels[percent_missing < 10] = ""
# Change point shapes and colors
p <- ggplot(data = data.frame(Mouse=seq_along(percent_missing),  
                         Percent_missing_genotype_data = percent_missing,
                         batch = factor(as.character(do.call(rbind.data.frame, strsplit(as.character(samples$Unique.Sample.ID), "_"))[,6]))
                         #batch = factor(as.character(do.call(rbind.data.frame, strsplit(as.character(samples$Directory), "_"))[,5]))
                         ), 
        aes(x=Mouse, y=Percent_missing_genotype_data, color = batch)) +
        geom_point() +
        geom_hline(yintercept=10, linetype="solid", color = "red") +
        geom_text_repel(aes(label=labels), vjust = 0, nudge_y = 0.01, show.legend = FALSE, size=3) +
        theme(text = element_text(size = 10))
p

dev.off()

p

save(percent_missing,file = "data/percent_missing_id_7.batches_myo.RData")

gm.covar = data.frame(id=rownames(gm$covar),gm$covar)
qc_info_cr <- merge(gm.covar,
                  data.frame(id = names(percent_missing),percent_missing = percent_missing,stringsAsFactors = F),by = "id")
bad.sample.cr <- qc_info_cr[qc_info_cr$percent_missing >= 10,]

```


```{r high missing table, echo=FALSE}

rownames(bad.sample.cr) <- NULL
bad.sample.cr[] <- lapply(bad.sample.cr, as.character)
bad.sample.cr$Sample_ID <- sapply(strsplit(as.character(bad.sample.cr$id),'_',fixed=T),function(x) x[7])
bad.sample.cr$Unique_ID <- bad.sample.cr$id
bad.sample.cr$batch.date <- ymd(parse_date_time((sapply(strsplit(as.character(bad.sample.cr$id),'_',fixed=T),function(x) x[6])), "ymd"))

bad.sample.cr[c("Sample_ID","Unique_ID","batch.date","percent_missing")] %>%
  kable(escape = F,align = c("ccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) #%>%
  #column_spec(1:3, width = "5cm") 

```

```{r removing bad samples CR, include=FALSE}

##removing bad samples
#gm <- gm[paste0("-",as.character(bad.sample.cr$id)),]#

#gm

```

### Sex
```{r Sexes, fig.height = 6, fig.width = 9.5, fig.align = "center"}
hdf5_filename <- dir(path = filepaths, pattern = "^hdf5_*", full.names = TRUE)
snps_file <- "/Users/corneb/Documents/MyJax/CS/Projects/support.files/MUGAarrays/UWisc/gm_uwisc_v1.csv"
snps <- read.csv(snps_file)

snps <- snps[snps$unique == TRUE, ]
#snps <- snps[snps$chr %in% c(1:19, "X"), ]
snps$chr <- sub("^chr", "", snps$chr)  ###remove prefix "chr"
colnames(snps)[colnames(snps)=="bp_mm10"] <- "pos" 
colnames(snps)[colnames(snps)=="cM_cox"] <- "cM"
snps <- snps %>% drop_na(chr, marker) 
snps$pos <- snps$pos * 1e-6
rownames(snps) <- snps$marker
colnames(snps)[1:4] <- c("marker", "chr", "pos", "pos") 

#  g <- h5read(hdf5_filename, "G")
#  g <- do.call(cbind, g)
x <- h5read(hdf5_filename, "X") # X channel intensities
x <- do.call(cbind, x)
#colnames(x) <- as.character(ind_ids(gm))
y <- h5read(hdf5_filename, "Y") # Y channel intensities
y <- do.call(cbind, y)
#colnames(y) <- as.character(ind_ids(gm))
rn <- h5read(hdf5_filename, "rownames")[[1]]  # markers 
cn <- h5read(hdf5_filename, "colnames")  # samples
cn <- do.call(c, cn)
# dimnames(g) <- list(rn, cn)
dimnames(x) <- list(rn, cn)
dimnames(y) <- list(rn, cn)
#cr <- colMeans(g != "--") # Call rate for each sample avg 0.95
#  sex <- determine_sex(x = x, y = y, markers = snps)$se

markers <- snps

chrx <- markers$marker[which(markers$chr == "X")]
chry <- markers$marker[which(markers$chr == "Y")]
#x[chrx,ind_ids(gm)]

chrx_int <- colMeans(x[chrx,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])] + y[chrx,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])], na.rm = T)
chry_int <- colMeans(x[chry,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])] + y[chry,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])], na.rm = T)

all.equal(as.character(ind_ids(gm)), as.character(samples$Unique.Sample.ID))

#sex order
#samples$Sex <- 'F'
sex <- samples$Sex


point_colors <- as.character( brocolors("web")[c("green", "purple")] )
percent_missing <- n_missing(gm, summary="proportion")*100
labels <- paste0(names(chrx_int), " (", round(percent_missing), "%)")
iplot( chrx_int,  chry_int, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Average Y chr intensity"))
```

For figures above and below, those labelled as female in metadata given, are coloured `green`, with those labelled as male are coloured as `purple`. The above is an interactive scatterplot of the average SNP intensity on the Y chromosome versus the average SNP intensity on the X chromosome.


```{r Sexes p2, fig.height = 6, fig.width = 9.5, fig.align = "center"}

phetX <- rowSums(gm$geno$X == 2)/rowSums(gm$geno$X != 0)
phetX <- phetX[as.character(do.call(rbind.data.frame, strsplit(names(phetX), "_"))[,7]) %in% names(chrx_int)]
names(phetX) <- as.character(do.call(rbind.data.frame, strsplit(names(phetX), "_"))[,7])
iplot(chrx_int, phetX, group=sex, indID=labels,
      chartOpts=list(pointcolor=point_colors, pointsize=4,
                     xlab="Average X chr intensity", ylab="Proportion het on X chr"))
```

In the above scatterplot, we show the proportion of hets vs the average intensity for the X chromosome SNPs. In calculating the proportion of heterozygous genotypes for the individuals, we look at X chromosome genotypes equal to 2 which corresponds to the heterozygote) relative to not being 0 (which is used to encode missing genotypes). The genotypes are arranged with rows being individuals and columns being markers. 

The following are the mice that have had sex incorrectly assigned:

```{r sex diff table, echo=FALSE, include=TRUE}

bad.sample.sex <- subset(samples, (samples$Inferred.Sex != samples$Sex))
bad.sample.sex$Sex[is.na(bad.sample.sex$Sex)] <- '--'
bad.sample.sex$Sample_ID <- sapply(strsplit(as.character(bad.sample.sex$Unique.Sample.ID),'_',fixed=T),function(x) x[7])
bad.sample.sex$Unique_ID <- bad.sample.sex$Unique.Sample.ID
bad.sample.sex$batch.date <- ymd(parse_date_time((sapply(strsplit(as.character(bad.sample.sex$Unique.Sample.ID),'_',fixed=T),function(x) x[6])), "ymd"))
rownames(bad.sample.sex) <- NULL

bad.sample.sex[c("Sample_ID","Unique_ID","batch.date","Sex","Inferred.Sex")] %>%
  kable(escape = F,align = c("ccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(1:4, width = "5cm") 

```

### Sample Duplicates
```{r Sample duplicates}

cg <- compare_geno(gm, cores=10)
summary.cg <- summary(cg)
```
```{r Sample duplicates table, echo=FALSE, include=FALSE}
rownames(summary.cg) <- NULL
summary.cg$iid1 <- summary.cg$ind1
summary.cg$iid2 <- summary.cg$ind2
summary.cg$pc.miss.ind1 <- percent_missing[match(summary.cg$ind1, names(percent_missing))]
summary.cg$pc.miss.ind2 <- percent_missing[match(summary.cg$ind2, names(percent_missing))]

#summary.cg$iid1 <- summary.cg$ind1
#summary.cg$iid2 <- summary.cg$ind2

#summary.cg[c(9,1,11,10,2,12,3:6)] %>%
#  kable(escape = F,align = c("llcccccc"),linesep ="\\hline") %>%
#  kable_styling(full_width = F) %>%
#  kable_styling("striped", full_width = F) %>%
#  column_spec(1:8, width = "5cm") 

```

Here is a histogram of the proportion of matching genotypes. The tick marks below the histogram indicate individual pairs.

```{r Sample duplicates figures}
save(summary.cg,file = "data/summary.cg_7.batches_myo.RData")

pdf(file = "output/Proportion_matching_genotypes_before_removal_of_bad_samples_7.batches_myo.pdf", width = 20, height = 20) 
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cg[upper.tri(cg)])
dev.off()

par(mar=c(5.1,0.6,0.6, 0.6))
hist(cg[upper.tri(cg)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes")
rug(cg[upper.tri(cg)])

```

```{r Sample duplicates 2 figures, include=TRUE}

cgsub <- cg[percent_missing < 10, percent_missing < 10]
par(mar=c(5.1,0.6,0.6, 0.6))
hist(cgsub[upper.tri(cgsub)], breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="Proportion matching genotypes [percent missing < 10%]")
rug(cgsub[upper.tri(cgsub)])

```

```{r top 20 high missing rate, include = FALSE}
##show top 20 samples with missing genotypes
#percent_missing <- n_missing(gm, "ind", "prop")*100
#round(sort(percent_missing, decreasing=TRUE)[1:20], 1)  
```

```{r removing duplicate samples, include=FALSE}
#removing duplcate pairs

#pheno_file_nick <- dir(path = filepathp, pattern = "*_pheno.csv+$", full.names = TRUE)
#pheno.pi <- read.csv(pheno_file_nick)

#master <- as.data.frame(read_excel("/Users/corneb/Documents/MyJax/CS/Projects/Simon.John.Tolman/Data/meta.data/inventory_file.xlsx", sheet = "inventory_file"))

#sample_file <- dir(path = filepaths, pattern = "^DODB_*", full.names = TRUE)
#samples <- read.csv(sample_file)

#project <- read.csv("/Users/corneb/Documents/MyJax/CS/Projects/Simon.John.Tolman/haplo.reconst/2020-11-09/all.samples/all.samples/project.csv", as.is=T)
#project$dir2 <- sapply(strsplit(project$dir,'/',fixed=T),function(x) x[7])

#inventory <- master[master$"Sample ID...4" %in% samples$Original.Mouse.ID,]
#invent.dir2 <- merge(inventory,project, by.x=c("Directory"), by.y=c("dir2"), all=T)
#invent.dodb <- merge(invent.dir2,samples, by.x=c("Directory","Sample ID...4"), by.y=c("Directory","Original.Mouse.ID"),all=T)

#gm.pheno1 <- merge(invent.dodb,pheno.pi, by.x=c("Sample ID...5"), by.y=c("Mouse.ID"), all.x=T)
#gm.pheno <- gm.pheno1[c("Correct.Mouse.ID","Unique.Sample.ID","Unique Sample ID","Sample ID...4","Sample ID...5","Directory",
#                     "Sex.x","Sex.y","Sex","Inferred.Sex","Birth Date","DOB","DO Generation","DO.Mother.Generation","DO.Generation","Strain.x","Strain.y",
#                     "Lmx1b allele","Lmx1b.allele","Lmx1b Genotype","Lmx1b.Genotype","Parental.Strain.GigaMUGA.ID","Parental Strain GigaMUGA ID",
#                     "Run.Date","Geneseek.Sample.ID","Include","cr","Max_IOP")]

#summary.cg$Name.ind1 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind1), "_"))[,6])
#summary.cg$Name.ind2 <- as.character(do.call(rbind.data.frame, strsplit(as.character(summary.cg$ind2), "_"))[,6])
summary.cg$Name.ind1 <- summary.cg$ind1
summary.cg$Name.ind2 <- summary.cg$ind2
summary.cg$miss.ind1 <- percent_missing[match(summary.cg$ind1, names(percent_missing))]
summary.cg$miss.ind2 <- percent_missing[match(summary.cg$ind2, names(percent_missing))]
#summary.cg$phen.ind1 <- gm.pheno$Max_IOP[match(summary.cg$ind1, gm.pheno$Unique.Sample.ID)]
#summary.cg$phen.ind2 <- gm.pheno$Max_IOP[match(summary.cg$ind2, gm.pheno$Unique.Sample.ID)]
summary.cg$remove.id <- ifelse((summary.cg$miss.ind1 > summary.cg$miss.ind2), summary.cg$ind1, summary.cg$ind2)
#summary.cg$remove.id.p1 <- ifelse(is.na(summary.cg$phen.ind1) & (is.na(summary.cg$phen.ind1) | !is.na(summary.cg$phen.ind1)), summary.cg$ind1, '')
#summary.cg$remove.id.p2 <- ifelse(is.na(summary.cg$phen.ind2) & (is.na(summary.cg$phen.ind1) | !is.na(summary.cg$phen.ind1)), summary.cg$ind2, '')
#remove.IDs <- rbind(summary.cg$remove.id.m[-1],summary.cg$remove.id.p1[-1],summary.cg$remove.id.p2[-1])
summary.cg$remove.id  

#qc_info$remove.id.duplicated <- ifelse(qc_info$id %in% summary.cg$remove.id, TRUE,FALSE)

#bad.sample.cr.dup <- NULL

#bad.sample.cr.dup$id <- c("Univ_of_Penn_Thaiss_MURMUGV01_20200130_9_A6",#
#                      "Univ_of_Penn_Thiass_MURGIGV01_20200226_90_C12", 
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_109_G2",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_111_B3",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_110_H5",
#                      "Univ_of_Penn_Thaiss_MURMUGV01_20200130_112_F5")

#gm <- gm[paste0("-",as.character(bad.sample.cr.dup$id)),]

#gm

#percent_missing <- n_missing(gm, "ind", "prop")*100
#round(sort(percent_missing, decreasing=TRUE)[1:19], 1)

```

### Array Intensities
```{r Array intensities, fig.height = 6, fig.width = 9.5, fig.align = "center"}

#load the intensities.fst_7.batches_myo.RData
#load("data/intensities.fst_7.batches_myo.RData")

xn <- x[,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])]
xn <- xn[snps$marker,]
xnm <- rownames(xn)

yn <- y[,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])]
yn <- yn[snps$marker,]

# bring together in one matrix
result <- cbind(snp=rep(snps$marker, 2),
                channel=rep(c("x", "y"), each=length(snps$marker)),
                as.data.frame(rbind(xn, yn)))
rownames(result) <- 1:nrow(result)

# bring SNP rows together
result <- result[as.numeric(t(cbind(seq_along(snps$marker), seq_along(snps$marker)+length(snps$marker)))),]
rownames(result) <- 1:nrow(result)

#load the intensities.fst_7.batches_myo.RData
#load("data/heh/intensities.fst_7.batches_myo.RData")
#X and Y channel
X <- result[result$channel == "x",]
rownames(X) <- X$snp
X <- X[,c(-1,-2)]

Y <- result[result$channel == "y",]
rownames(Y) <- Y$snp
Y <- Y[,c(-1,-2)]

int <- result

#int <- result

#rm(result)
int <- int[seq(1, nrow(int), by=2),-(1:2)] + int[-seq(1, nrow(int), by=2),-(1:2)]
names(int) <- gsub("\\.1","", names(int))
int <- int[,as.character(do.call(rbind.data.frame, strsplit(ind_ids(gm), "_"))[,7])]
names(int) <- gsub("\\.1","", names(int))
names(int) <- as.character(names(percent_missing))
names(percent_missing) <- as.character(names(percent_missing))
n <- names(sort(percent_missing[intersect(as.character(ind_ids(gm)), colnames(int))], decreasing=TRUE))
tn = int[,n]
iboxplot(log10(t(tn)+1), orderByMedian=FALSE, chartOpts=list(ylab="log10(SNP intensity + 1)"))
names(tn) <- as.character(do.call(rbind.data.frame, strsplit(names(tn), "_"))[,7])
iboxplot(log10(t(tn)+1), orderByMedian=FALSE, chartOpts=list(ylab="log10(SNP intensity + 1)"))

```
In the above plot, distributions of array intensities (after a log10(x+1) transformation) are displayed. 

The arrays are sorted by the proportion of missing genotype data for the sample, and the curves connect various quantiles of the intensities.

```{r Array intensities percentile figure, fig.height = 6, fig.width = 9.5, fig.align = "center"}

qu <- apply(int, 2, quantile, c(0.01, 0.99), na.rm=TRUE)
group <- (percent_missing >= 19.97) + (percent_missing > 5) + (percent_missing > 2) + 1
labels <- paste0(as.character(do.call(rbind.data.frame, strsplit(colnames(qu), "_"))[,7]), " (", round(percent_missing), "%)")
iplot(qu[1,], qu[2,], indID=labels, group=group,
      chartOpts=list(xlab="1 %ile of array intensities",
                     ylab="99 %ile of array intensities",
                     pointcolor=c("#ccc", "slateblue", "Orchid", "#ff851b")))
```
For this particular set of arrays, a plot of the 1 %ile vs the 99 %ile is quite revealing. In the following, the orange points are those with > 20% missing genotypes, the pink points are the samples with 5-20% missing genotypes, and the blue points are the samples with 2-5% missing genotypes.


### Genotyping Error LOD Scores
```{r  Genotyping error LOD scores, fig.height = 6, fig.width = 9.5, fig.align = "center"}
load("data/e_7.batches_myo.RData")
errors_ind <- rowSums(e>2)[rownames(gm$covar)]/n_typed(gm)*100
lab <- paste0(as.character(do.call(rbind.data.frame, strsplit(names(errors_ind), "_"))[,7]), " (", myround(percent_missing[as.character(rownames(gm$covar))],1), "%)")
iplot(seq_along(errors_ind), errors_ind, indID=lab,
      chartOpts=list(xlab="Mouse", ylab="Percent genotyping errors", ylim=c(0, 15),
                     axispos=list(xtitle=25, ytitle=50, xlabel=5, ylabel=5)))
save(errors_ind, file = "data/errors_ind_7.batches_myo.RData")

```

### Removing Samples
```{r removing samples }

##percent missing
gm.covar = data.frame(id=as.character(rownames(gm$covar)),gm$covar)
qc_info <- merge(gm.covar,
                  data.frame(id = names(percent_missing),percent_missing = percent_missing,stringsAsFactors = F),by = "id")

#missing sex
#qc_info$sex.match <- ifelse(qc_info$sexp == qc_info$sex, TRUE, FALSE)
rownames(samples) <- as.character(samples$Unique.Sample.ID)
samples <- samples[as.character(qc_info$id),]
#samples$Unique.Sample.ID <- as.character(samples$Unique.Sample.ID)
all.equal(as.character(qc_info$id), as.character(samples$Unique.Sample.ID))
qc_info$sex.match <- ifelse((samples$Inferred.Sex == samples$Sex), TRUE, FALSE)

#genotype errors
qc_info <- merge(qc_info,
                 data.frame(id = as.character(names(errors_ind)),
                            genotype_erros = errors_ind,stringsAsFactors = F),by = "id")

##duplicated id to be remove
qc_info$duplicate.id <- ifelse(qc_info$id %in% as.character(summary.cg$remove.id), TRUE,FALSE)

#bad.sample <- qc_info[qc_info$generation ==1 | qc_info$Number_crossovers <= 200 | qc_info$Number_crossovers >=1000 | qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 1 | qc_info$remove.id.duplicated == TRUE,]
bad.sample <- qc_info[qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 8,]

save(qc_info, bad.sample, file = "data/qc_info_bad_sample_7.batches_myo.RData")

gm_samqc <- gm[paste0("-",as.character(bad.sample$id)),]

gm_samqc

save(gm_samqc, file = "data/gm_samqc_7.batches_myo.RData")

# update other stuff
e <- e[ind_ids(gm_samqc),]
#g <- g[ind_ids(gm_samqc),]
#snpg <- snpg[ind_ids(gm_samqc),]

#save(e,g,snpg, file = "data/e_g_snpg_samqc_7.batches_myo.RData")
save(e, file = "data/e_snpg_samqc_7.batches_myo.RData")

```

Here is the list of samples that were removed:

```{r removing table, echo=FALSE}

removed.ids <- bad.sample[c(1)]
names(removed.ids) <- "Unique_ID"
removed.ids$Sample_ID <-  as.character(do.call(rbind.data.frame, strsplit(as.character(removed.ids$Unique_ID), "_"))[,7])
removed.ids$batch.date <- ymd(parse_date_time((sapply(strsplit(as.character(removed.ids$Unique_ID),'_',fixed=T),function(x) x[6])), "ymd"))

rownames(removed.ids) <- NULL

removed.ids[c("Sample_ID","Unique_ID","batch.date")] %>%
  kable(escape = F,align = c("c"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F) 

```

Below is a table summarising the problematic samples found throughout QC. These include the following:

+ no_pheno == sample does not have a phenotype (if applicable)
+ high_miss == sample has higher than 10% missing genotypes 
+ diff_sex == sample has sex mismatch
+ high_geno_errors == Sample has geno errors above 10%
+ dublicate_ID == sample was duplicated (or highly concordant)

**NB:** For duplcate pairs, the one that was chosen to be removed was the one that had a higher missing rate

```{r removing samples table, echo=FALSE}


#pheno_file_nick <- dir(path = filepathp, pattern = "*_pheno.csv+$", full.names = TRUE)
#gm.pheno <- read.csv(pheno_file_nick)
#gm.pheno$id <-  gm.pheno$Unique.Sample.ID
#qc_info <- merge(qc_info,gm.pheno,by = "id",all.x=T)

#bad.sample <- qc_info[qc_info$ngen ==1 | qc_info$Number_crossovers <= 200 | qc_info$Number_crossovers >=1000 | qc_info$percent_missing >= 10 | qc_info$genotype_erros >= 1 | qc_info$remove.id.duplicated == TRUE | is.na(qc_info$Time_min)==TRUE | is.na(qc_info$Distance_m)==TRUE | is.na(qc_info$Energy_J)==TRUE,]

ind1<-summary.cg[c("ind1","iid1")]
names(ind1) <- c("iid","id")
ind2<-summary.cg[c("ind2","iid2")]
names(ind2) <- c("iid","id")
dup.IDS <- rbind(ind1, ind2)
dup.IDS <-dup.IDS[!duplicated(dup.IDS),]
dup.IDS$duplicated <- "TRUE"

bad.sample <- merge(qc_info,dup.IDS,by=c("id"),all=T)
bad.sample[is.na(bad.sample)] <- ""

bad <- NULL
#bad$Mouse_ID <- bad.sample$"Sample ID...5"
#bad<-as.data.frame(bad)
#bad$Neogen_Sample_ID <- bad.sample$"Sample ID...4"
bad$Unique_ID <- as.character(bad.sample$id)
bad$Sample_ID <-  as.character(do.call(rbind.data.frame, strsplit(as.character(bad$Unique_ID), "_"))[,7])
bad$batch.date <- ymd(parse_date_time((sapply(strsplit(as.character(bad$Unique_ID),'_',fixed=T),function(x) x[6])), "ymd"))
#bad$Neogen_Sample_ID <- as.character(do.call(rbind.data.frame, strsplit(as.character(bad$Genoprobs_Sample_ID), "_"))[,7])
#bad$Unique_Sample_ID <- bad.sample$id
#bad$no_pheno <- ifelse(is.na(bad.sample$Max_IOP)==TRUE, 'XX', '')
bad$high_miss <- ifelse(bad.sample$percent_missing >= 10, 'XX', '')
bad$diff_sex <-  ifelse(bad.sample$sex.match ==FALSE, 'XX', '')
#bad$high_xo <- ifelse(bad.sample$Number_crossovers <= 200 | bad.sample$Number_crossovers >=1000, 'XX', '')
#bad$DO.Generation <- ifelse(bad.sample$generation ==1, 'XX', '')
bad$high_geno.errors <- ifelse(bad.sample$genotype_erros >= 10, 'XX', '')
bad$highly_concordant <- ifelse(bad.sample$duplicated == TRUE, 'XX', '')
bad[is.na(bad)] <- ""
bad[bad=='NA'] <- ""
df <- do.call(cbind, bad)
bad <- as.data.frame(df)

badind <- subset(bad, 
         #bad$no_pheno == 'XX'|
         bad$high_miss == 'XX'|
         bad$diff_sex == 'XX'|  
#         bad$high_xo == 'XX'|
         bad$high_geno.errors == 'XX'|
#         bad$DO.Generation == 'XX'|
         bad$highly_concordant == 'XX')


#badind <- bad[bad$no_pheno == 'XX',]

badind[] <- lapply(badind, as.character)
#badind$Thaiss_ID <- ifelse(badind$Thaiss == 994 | badind$Thaiss == 995 | badind$Thaiss == 996 |badind$Thaiss == 997 | badind$Thaiss == 998 | badind$Thaiss == 999, "--", bad$Thaiss_ID)

rownames(badind) <- NULL

badind[c(2,1,3:7)] %>% 
   mutate(
     #no_pheno = ifelse(no_pheno == 'XX',
     #             cell_spec(no_pheno, color = 'green',background = 'green'),
     #             ''),
     high_miss = ifelse(high_miss == 'XX',
                  cell_spec(high_miss, color = 'red',background = 'red'),
                  ''),
     diff_sex = ifelse(diff_sex == 'XX',
                  cell_spec(diff_sex, color = 'blue',background = 'blue'),
                  ''),
#     high_xo = ifelse(high_xo == 'XX',
#                  cell_spec(high_xo, color = 'purple',background = 'purple'),
#                  ''),
     high_geno.errors = ifelse(high_geno.errors == 'XX',
                  cell_spec(high_geno.errors, color = 'pink',background = 'pink'),
                  ''),
#     DO.Generation = ifelse(DO.Generation == 'XX',
#                  cell_spec(DO.Generation, color = 'lightgreen',background = 'lightgreen'),
#                  ''),
     highly_concordant = ifelse(highly_concordant == 'XX',
                  cell_spec(highly_concordant, color = 'orange',background = 'orange'),
                  '')
     ) %>%
   kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
   kable_styling("striped", full_width = F) %>%
   #column_spec(2:6, width = "3cm") %>%
   column_spec(1, width = "5cm")

```

