---
title: "Step 2: SNP QC [4.batches_myo]"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/qc/workflowr/Serreze-T1D_Workflow")

#libraries
library(broman)
library(qtl2)
library(qtlcharts)
library(ggplot2)
library(ggrepel)
#library(DOQTL)
library(mclust)
#source("code/reconst_utils.R")
library("kableExtra")
library("knitr")
library("fst")
```

### Loading Project
```{r loading project}
load("data/e_snpg_samqc_4.batches_myo.RData")
gm <- get(load("data/gm_samqc_4.batches_myo.RData"))

gm
```

It can also be useful to look at the proportion of missing genotypes by marker. Markers with a lot of missing data were likely difficult to call, and so the genotypes that were called may contain a lot of errors.

### Marker Missing Data 
```{r Missing data in Markers and Genotype frequencies Markers, fig.height = 6, fig.width = 9.5, fig.align = "center"}

pmis_mar <- n_missing(gm, "marker", "proportion")*100
save(pmis_mar, file = "data/percent_missing_marker_4.batches_myo.RData")

par(mar=c(5.1,0.6,0.6, 0.6))
hist(pmis_mar, breaks=seq(0, 100, length=201),
     main="", yaxt="n", ylab="", xlab="Percent missing genotypes")
rug(pmis_mar)

pdf(file = "output/Percent_missing_genotype_data_per_marker_4.batches_myo.pdf")
par(mar=c(5.1,0.6,0.6, 0.6))
hist(pmis_mar, breaks=seq(0, 100, length=201),
     main="", yaxt="n", ylab="", xlab="Percent missing genotypes")
rug(pmis_mar)
dev.off()
```

```{r missing table, echo=FALSE}

pmis <- NULL
pmis$pmis_mar_5 <- sum(pmis_mar >= 5)
pmis$pmis_mar_10 <- sum(pmis_mar >= 10)
pmis$pmis_mar_15 <- sum(pmis_mar >= 15)
pmis$pmis_mar_25 <- sum(pmis_mar >= 25)
pmis$pmis_mar_50 <- sum(pmis_mar >= 50)
pmis$pmis_mar_75 <- sum(pmis_mar >= 75)
pmis$total_snps <- nrow(as.data.frame(pmis_mar))

pmis <- t(as.data.frame(pmis))
pmis <- as.data.frame(pmis)
pmis$count <- pmis$V1

pmis[c(2)] %>%
  kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F)  %>%
  row_spec(7 ,bold=T,color= "white",background = "black")

```

### Marker Genotype Frequencies
```{r Genotype frequencies Markers, fig.height = 6, fig.width = 9.5, fig.align = "center"}

g <- do.call("cbind", gm$geno[1:19])
#fg <- do.call("cbind", gm$founder_geno[1:19])
#g <- g[,colSums(g)!=0]
#fg <- fg[,colSums(fg==0)==0]
#fgn <- colSums(g==2)

gf_mar <- t(apply(g, 2, function(a) table(factor(a, 1:2))/sum(a != 0)))
gn_mar <- t(apply(g, 2, function(a) table(factor(a, 1:2))))

gf_mar <- gf_mar[gf_mar[,2] != "NaN",]
MAF <- apply(gf_mar, 1, function(x) min(x))
MAF <- as.data.frame(MAF)
MAF$index <- 1:nrow(gf_mar)
gf_mar_maf <- merge(gf_mar,as.data.frame(MAF), by="row.names")
gf_mar_maf <- gf_mar_maf[order(gf_mar_maf$index),]

pdf(file = "output/genotype_frequency_marker_4.batches_myo.pdf")
par(mar=c(5.1,0.6,0.6, 0.6))
hist(gf_mar_maf$MAF, breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="MAF")
rug(gf_mar_maf$MAF)
dev.off()

par(mar=c(5.1,0.6,0.6, 0.6))
hist(gf_mar_maf$MAF, breaks=seq(0, 1, length=201),
     main="", yaxt="n", ylab="", xlab="MAF")
rug(gf_mar_maf$MAF)

gfmar <- NULL
gfmar$gfmar_mar_0 <- sum(gf_mar_maf$MAF==0)
gfmar$gfmar_mar_1 <- sum(gf_mar_maf$MAF< 0.01)
gfmar$gfmar_mar_5 <- sum(gf_mar_maf$MAF< 0.05)
gfmar$gfmar_mar_10 <- sum(gf_mar_maf$MAF< 0.10)
gfmar$gfmar_mar_15 <- sum(gf_mar_maf$MAF< 0.15)
gfmar$gfmar_mar_25 <- sum(gf_mar_maf$MAF< 0.25)
gfmar$gfmar_mar_50 <- sum(gf_mar_maf$MAF<= 0.50)
gfmar$total_snps <- nrow(as.data.frame(gf_mar_maf))

gfmar <- t(as.data.frame(gfmar))
gfmar <- as.data.frame(gfmar)
gfmar$count <- gfmar$V1

gfmar[c(2)] %>%
  kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
  kable_styling(full_width = F) %>%
  kable_styling("striped", full_width = F)  %>%
  row_spec(8 ,bold=T,color= "white",background = "black")

save(gf_mar, file = "data/genotype_freq_marker_4.batches_myo.RData")

```

### Marker Genotype Errors
```{r Genotype errors Markers, fig.height = 6, fig.width = 9.5, fig.align = "center"}
errors_mar <- colSums(e>2)/n_typed(gm, "marker")*100

grayplot(pmis_mar, errors_mar,
         xlab="Proportion missing", ylab="Proportion genotyping errors")


pdf(file = "output/genotype_error_marker_4.batches_myo.pdf")
grayplot(pmis_mar, errors_mar,
         xlab="Proportion missing", ylab="Proportion genotyping errors")
dev.off()

save(errors_mar, file = "data/genotype_errors_marker_4.batches_myo.RData")

```

Markers with higher rates of missing genotypes tend to show higher errors rates.

### Removing Markers
#### Missingness
```{r listing markers missing}

#Missingness

length(pmis_mar[pmis_mar >= 10])

high_miss <- find_markerpos(gm, names(pmis_mar[pmis_mar >= 10]))
high_miss$id <- rownames(high_miss)
high_miss_df <- as.data.frame(pmis_mar[pmis_mar >= 10])
high_miss_df$index = 1: nrow(high_miss_df)
high_miss_df$id <- rownames(high_miss_df)

high_miss_bad <- merge(high_miss,high_miss_df, by=c("id"),all=T)
names(high_miss_bad)[5] <- c("high_miss")
names(high_miss_bad)[1] <- c("marker")
high_miss_bad <- high_miss_bad[order(high_miss_bad$index),]
```

#### Monomorphic/Low Frequency markers
```{r listing markers Monomorphic or Low Frequency markers}

#Monomorphic/Low Frequency markers

#low_freq_df <- as.data.frame(gf_mar)
count <- rowSums(gf_mar <= 0.01)
#count <- as.data.frame(count)
low_freq_df <- merge(as.data.frame(gf_mar),as.data.frame(count), by="row.names",all=T)
low_freq_df[is.na(low_freq_df)] <- ''
low_freq_df <- low_freq_df[low_freq_df$count == 1,]
rownames(low_freq_df) <- low_freq_df$Row.names
#low_freq_df$id <- rownames(low_freq_df)
#low_freq_df$index = 1: nrow(low_freq_df)
low_freq <- find_markerpos(gm, rownames(low_freq_df))
low_freq$id <- rownames(low_freq)

nrow(low_freq)

low_freq_bad <- merge(low_freq,low_freq_df, by="row.names",all=T)
#names(low_freq_bad)[5] <- c("AA_freq")
#names(low_freq_bad)[6] <- c("AB_freq")
#names(low_freq_bad)[7] <- c("BB_freq")
names(low_freq_bad)[1] <- c("marker")
#low_freq_bad <- low_freq_bad[order(low_freq_bad$index),]

```

#### Genotyping Error
```{r listing markers Genotyping Error}

##Genotyping Error

length(errors_mar[errors_mar > 5])

error_markers_names <- names(errors_mar[errors_mar > 5])
error_markers_names <- error_markers_names[complete.cases(error_markers_names)]

error_markers <- find_markerpos(gm, error_markers_names)
error_markers$id <- rownames(error_markers)
#rne <- rownames(as.data.frame(errors_mar))
error_mars_df <- as.data.frame(errors_mar[errors_mar > 5])
error_mars_df <- error_mars_df[complete.cases(error_mars_df$"errors_mar[errors_mar > 5]"),]
error_mars_df <- as.data.frame(error_mars_df)
#error_mars_df$id = rownames(error_mars_df)
error_mars_df$index = 1: nrow(error_mars_df)

#error_markers_bad <- merge(error_markers,error_mars_df, by=c("id"),all=T)
error_markers_bad <- cbind(error_markers,error_mars_df)
names(error_markers_bad)[5] <- c("error_mars")
names(error_markers_bad)[4] <- c("marker")
error_markers_bad <- error_markers_bad[order(error_markers_bad$index),]

```

#### Total
```{r listing markers total}

### merge all

bad_markers <- rbind(high_miss_bad[c("marker","chr","gmap","pmap")], low_freq_bad[c("marker","chr","gmap","pmap")], error_markers_bad[c("marker","chr","gmap","pmap")])
#nrow(bad_markers)

duplicate <- bad_markers[duplicated(bad_markers),]

bad_markers <- bad_markers[!duplicated(bad_markers),]
nrow(bad_markers)

save(bad_markers, file = "data/bad_markers_all_4.batches_myo.RData")

```

Only removing markers that are missing in at least 10% of the samples

```{r Removing markers }

#missing in at least 10% of the samples

gm_allqc2 <- drop_markers(gm_samqc, high_miss_bad$marker)
gm_allqc <- drop_nullmarkers(gm_allqc2)


gm_allqc

save(gm_allqc, file = "data/gm_allqc_4.batches_myo_mis.RData")

```
