---
title: "Phenotype QC [Continious Phenotype - Age of Onset]"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)

library("kableExtra")
library("knitr")
options(stringsAsFactors = FALSE)
library(data.table) 
library(tidyr)     
library(mclust)     
library(rhdf5)      
library(optparse)
library(dplyr)
library(cluster)
library(readxl)
library(psych)
library(tibble)
library(ggplot2)
library(reshape2)
library(qtl2)
#library(qtl)
library(abind)

setwd("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/qc/workflowr/Serreze-T1D_Workflow")

```

### Loading Data

```{r data, echo=TRUE}

load("data/gm_allqc_4.batches.RData")

#gm_allqc
gm=gm_allqc
gm

pr <- readRDS("data/serreze_probs_allqc.rds")
#pr <- readRDS("data/serreze_probs.rds")

```

### ICI vs EOI

```{r Phenotype distribution1, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}


##extracting animals with ici and eoi group status
miceinfo <- gm$covar[gm$covar$group == "EOI" | gm$covar$group == "ICI",]
table(miceinfo$group)
mice.ids <- rownames(miceinfo)

gm <- gm[mice.ids]
gm

pr.qc <- pr
for (i in 1:20){pr.qc[[i]] = pr.qc[[i]][mice.ids,,]}

gm$covar$ICI.vs.EOI <- ifelse(gm$covar$group == "EOI", 0, 1)
names(gm$covar)[3] <- c("age.of.onset")
gm$covar$age.of.onset <- as.numeric(gm$covar$age.of.onset)

p <- ggplot(gm$covar, aes(x=as.numeric(age.of.onset))) + geom_histogram(color="black", fill="white")
p

phenos.covars.lg <- gm$covar %>% gather(variable, value, -c("id","group","sex","diabetic status","strain","ICI.vs.EOI"))

box1 <-  ggplot(data=phenos.covars.lg, aes(x=variable, y=as.numeric(value), color=group, fill=group)) +
         geom_boxplot(position = position_dodge(width=0.9)) +
         #ggtitle(paste0("Values of ",v," [random dataframe: ",r,"]")) +
         #labs(y = v) +
         theme(strip.text.x = element_text(size=13),
               axis.text.x = element_text(size = 13, angle = 0),
               axis.text.y = element_text(size = 13, angle = 0),  
               axis.title.x=element_blank(),
               axis.title.y=element_blank(),
               plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
                  #legend.position = "none"
             )
box1


```
QTL analysis requires variables follow normal distribution, from the above distributions, we need to ranknorm the data.  


```{r Phenotype distribution2, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}


##ranknorm
rz.transform <- function(y) {
  rankY=rank(y, ties.method="average", na.last="keep")
  rzT=qnorm(rankY/(length(na.exclude(rankY))+1))
  return(rzT)
}

gm$covar$rz.age <- rz.transform(gm$covar$age.of.onset)

p <- ggplot(gm$covar, aes(x=as.numeric(rz.age))) + geom_histogram(color="black", fill="white")
p

phenos.covars.lg <- gm$covar %>% gather(variable, value, -c("id","group","sex","diabetic status","strain","ICI.vs.EOI","age.of.onset"))

box1 <-  ggplot(data=phenos.covars.lg, aes(x=variable, y=as.numeric(value), color=group, fill=group)) +
         geom_boxplot(position = position_dodge(width=0.9)) +
         #ggtitle(paste0("Values of ",v," [random dataframe: ",r,"]")) +
         #labs(y = v) +
         theme(strip.text.x = element_text(size=13),
               axis.text.x = element_text(size = 13, angle = 0),
               axis.text.y = element_text(size = 13, angle = 0),  
               axis.title.x=element_blank(),
               axis.title.y=element_blank(),
               plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
                  #legend.position = "none"
             )
box1


```

And then remove any samples that are 3 standard deviations from the mean. 

```{r outliers, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}


#outliers
#des.3 <- Hmisc::describe(df_phenos[,c("R_AVG","L_AVG","Both_AVG")]) 
des.1 <- pastecs::stat.desc(gm$covar[,c("age.of.onset" ,"rz.age")]) 
des.2 <- psych::describe(gm$covar[,c("age.of.onset" ,"rz.age")]) 

#scale(df_phenos[,c("R_AVG")])

gm$covar$out.age.of.onset <- ifelse(gm$covar[,c("age.of.onset")] > (des.1[9,1] + 3*des.1[13,1])  | gm$covar[,c("age.of.onset")] < (des.1[9,1] - 3*des.1[13,1]), 'Outlier','Keep')
gm$covar$out.rz.age <- ifelse(gm$covar[,c("rz.age")] > (des.1[9,2] + 3*des.1[13,2])  | gm$covar[,c("rz.age")] < (des.1[9,2] - 3*des.1[13,2]), 'Outlier','Keep')

bad <- NULL
bad$Mouse.ID <- rownames(gm$covar)
bad$age.of.onset <- ifelse(gm$covar$out.age.of.onset =="Outlier", 'XX', '')
bad$rz.age <- ifelse(gm$covar$out.rz.age =="Outlier", 'XX', '')
bad[is.na(bad)] <- ""
bad[bad=='NA'] <- ""
df <- do.call(cbind, bad)
bad <- as.data.frame(df)

badind <- subset(bad, 
         bad$age.of.onset == 'XX'|
         bad$rz.age == 'XX')


#badind <- bad[bad$no_pheno == 'XX',]

badind[] <- lapply(badind, as.character)
#badind$Thaiss_ID <- ifelse(badind$Thaiss == 994 | badind$Thaiss == 995 | badind$Thaiss == 996 |badind$Thaiss == 997 | badind$Thaiss == 998 | badind$Thaiss == 999, "--", bad$Thaiss_ID)

rownames(badind) <- NULL

badind[] %>% 
   dplyr::mutate(
     age.of.onset = ifelse(age.of.onset == 'XX',
                  cell_spec(age.of.onset, color = 'gray',background = 'gray'),
                  ''),
     rz.age = ifelse(rz.age == 'XX',
                  cell_spec(rz.age, color = 'gray',background = 'gray'),
                  '')
     ) %>%
   kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
   kable_styling("striped", full_width = F) %>%
   column_spec(1:3, width = "3cm") 



##removing outliers
gm$covar$Mouse.ID <- rownames(gm$covar)
gm$covar$age.of.onset[gm$covar$out.age.of.onset == "Outlier"] <- '' 
gm$covar$rz.age[gm$covar$out.rz.age == "Outlier"] <- '' 

#gm$covar <- gm$covar[c(1:15)]
#gm$covar$id <- rownames(gm$covar)
write.csv(gm$covar,"data/covar_cleaned_ici.vs.eoi.csv", row.names=F, quote=F)

```

That is, those that have a grey square were removed for that particular phenotype in the QTL mapping. 


### ICI vs PBS

```{r Phenotype distribution21, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}

gm=gm_allqc

##extracting animals with ici and eoi group status
miceinfo <- gm$covar[gm$covar$group == "PBS" | gm$covar$group == "ICI",]
table(miceinfo$group)
mice.ids <- rownames(miceinfo)

gm <- gm[mice.ids]
gm

pr.qc <- pr
for (i in 1:20){pr.qc[[i]] = pr.qc[[i]][mice.ids,,]}

gm$covar$ICI.vs.PBS <- ifelse(gm$covar$group == "PBS", 0, 1)
names(gm$covar)[3] <- c("age.of.onset")
gm$covar$age.of.onset <- as.numeric(gm$covar$age.of.onset)

p <- ggplot(gm$covar, aes(x=as.numeric(age.of.onset))) + geom_histogram(color="black", fill="white")
p

phenos.covars.lg <- gm$covar %>% gather(variable, value, -c("id","group","sex","diabetic status","strain","ICI.vs.PBS"))

box1 <-  ggplot(data=phenos.covars.lg, aes(x=variable, y=as.numeric(value), color=group, fill=group)) +
         geom_boxplot(position = position_dodge(width=0.9)) +
         #ggtitle(paste0("Values of ",v," [random dataframe: ",r,"]")) +
         #labs(y = v) +
         theme(strip.text.x = element_text(size=13),
               axis.text.x = element_text(size = 13, angle = 0),
               axis.text.y = element_text(size = 13, angle = 0),  
               axis.title.x=element_blank(),
               axis.title.y=element_blank(),
               plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
                  #legend.position = "none"
             )
box1


```
QTL analysis requires variables follow normal distribution, from the above distributions, we need to ranknorm the data.  


```{r Phenotype distribution22, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}
##ranknorm
rz.transform <- function(y) {
  rankY=rank(y, ties.method="average", na.last="keep")
  rzT=qnorm(rankY/(length(na.exclude(rankY))+1))
  return(rzT)
}

gm$covar$rz.age <- rz.transform(gm$covar$age.of.onset)

p <- ggplot(gm$covar, aes(x=as.numeric(rz.age))) + geom_histogram(color="black", fill="white")
p

phenos.covars.lg <- gm$covar %>% gather(variable, value, -c("id","group","sex","diabetic status","strain","ICI.vs.PBS","age.of.onset"))

box1 <-  ggplot(data=phenos.covars.lg, aes(x=variable, y=as.numeric(value), color=group, fill=group)) +
         geom_boxplot(position = position_dodge(width=0.9)) +
         #ggtitle(paste0("Values of ",v," [random dataframe: ",r,"]")) +
         #labs(y = v) +
         theme(strip.text.x = element_text(size=13),
               axis.text.x = element_text(size = 13, angle = 0),
               axis.text.y = element_text(size = 13, angle = 0),  
               axis.title.x=element_blank(),
               axis.title.y=element_blank(),
               plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
                  #legend.position = "none"
             )
box1


```

And then remove any samples that are 3 standard deviations from the mean. 

```{r outliers2, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE}

#outliers
#des.3 <- Hmisc::describe(df_phenos[,c("R_AVG","L_AVG","Both_AVG")]) 
des.1 <- pastecs::stat.desc(gm$covar[,c("age.of.onset" ,"rz.age")]) 
des.2 <- psych::describe(gm$covar[,c("age.of.onset" ,"rz.age")]) 

#scale(df_phenos[,c("R_AVG")])

gm$covar$out.age.of.onset <- ifelse(gm$covar[,c("age.of.onset")] > (des.1[9,1] + 3*des.1[13,1])  | gm$covar[,c("age.of.onset")] < (des.1[9,1] - 3*des.1[13,1]), 'Outlier','Keep')
gm$covar$out.rz.age <- ifelse(gm$covar[,c("rz.age")] > (des.1[9,2] + 3*des.1[13,2])  | gm$covar[,c("rz.age")] < (des.1[9,2] - 3*des.1[13,2]), 'Outlier','Keep')

bad <- NULL
bad$Mouse.ID <- rownames(gm$covar)
bad$age.of.onset <- ifelse(gm$covar$out.age.of.onset =="Outlier", 'XX', '')
bad$rz.age <- ifelse(gm$covar$out.rz.age =="Outlier", 'XX', '')
bad[is.na(bad)] <- ""
bad[bad=='NA'] <- ""
df <- do.call(cbind, bad)
bad <- as.data.frame(df)

badind <- subset(bad, 
         bad$age.of.onset == 'XX'|
         bad$rz.age == 'XX')


#badind <- bad[bad$no_pheno == 'XX',]

badind[] <- lapply(badind, as.character)
#badind$Thaiss_ID <- ifelse(badind$Thaiss == 994 | badind$Thaiss == 995 | badind$Thaiss == 996 |badind$Thaiss == 997 | badind$Thaiss == 998 | badind$Thaiss == 999, "--", bad$Thaiss_ID)

rownames(badind) <- NULL

badind[] %>% 
   dplyr::mutate(
     age.of.onset = ifelse(age.of.onset == 'XX',
                  cell_spec(age.of.onset, color = 'gray',background = 'gray'),
                  ''),
     rz.age = ifelse(rz.age == 'XX',
                  cell_spec(rz.age, color = 'gray',background = 'gray'),
                  '')
     ) %>%
   kable(escape = F,align = c("ccccccccc"),linesep ="\\hline") %>%
   kable_styling("striped", full_width = F) %>%
   column_spec(1:3, width = "3cm") 



##removing outliers
gm$covar$Mouse.ID <- rownames(gm$covar)
gm$covar$age.of.onset[gm$covar$out.age.of.onset == "Outlier"] <- '' 
gm$covar$rz.age[gm$covar$out.rz.age == "Outlier"] <- '' 

#gm$covar <- gm$covar[c(1:15)]
#gm$covar$id <- rownames(gm$covar)
write.csv(gm$covar,"data/covar_cleaned_ici.vs.pbs.csv", row.names=F, quote=F)

```

That is, those that have a grey square were removed for that particular phenotype in the QTL mapping. 