---
title: "QTL Analysis - binary [ICI vs EOI]"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)

library("kableExtra")
library("knitr")
options(stringsAsFactors = FALSE)
library(data.table) 
library(tidyr)     
library(mclust)     
library(rhdf5)      
library(optparse)
library(dplyr)
library(cluster)
library(readxl)
library(psych)
library(tibble)
library(ggplot2)
library(reshape2)
library(qtl2)
#library(qtl)
library(abind)

setwd("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/qc/workflowr/Serreze-T1D_Workflow")

```

### Loading Data

We will load the data and subset indivials out that are in the groups of interest.  We will create a binary phenotype from this  (EOI ==0, ICI == 1).

```{r data, echo=TRUE, results="asis"}

load("data/gm_allqc_4.batches.RData")

#gm_allqc
gm=gm_allqc
gm

pr <- readRDS("data/serreze_probs_allqc.rds")
#pr <- readRDS("data/serreze_probs.rds")

##extracting animals with ici and eoi group status
miceinfo <- gm$covar[gm$covar$group == "EOI" | gm$covar$group == "ICI",]
table(miceinfo$group)
mice.ids <- rownames(miceinfo)

gm <- gm[mice.ids]
gm

pr.qc <- pr
for (i in 1:20){pr.qc[[i]] = pr.qc[[i]][mice.ids,,]}

#bin_pheno <- NULL
#bin_pheno$EOI <- ifelse(gm$covar$group == "EOI", 1, 0)
#bin_pheno$ICI <- ifelse(gm$covar$group == "ICI", 1, 0)
#bin_pheno <- as.data.frame(bin_pheno)
#rownames(bin_pheno) <- rownames(gm$covar)

gm$covar$ICI.vs.EOI <- ifelse(gm$covar$group == "EOI", 0, 1)

#pmap_interp = interp_map(map, gm$gmap, gm$pmap)
markers <- marker_names(gm)
gmapdf <- read.csv("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output_hh/genetic_map.csv")
pmapdf <- read.csv("/Users/corneb/Documents/MyJax/CS/Projects/Serreze/haplotype.reconstruction/output_hh/physical_map.csv")
mapdf <- merge(gmapdf,pmapdf, by=c("marker","chr"), all=T)
rownames(mapdf) <- mapdf$marker
#mapdf$chr = sapply(strsplit(mapdf$marker,'_'),function(x) x[1])
#mapdf$chr = sub("^c","", mapdf$chr)
#mapdf$chr = sapply(strsplit(mapdf$chr,'.loc'),function(x) x[1])
mapdf <- mapdf[markers,]
#mapdf <- mapdf[order(mapdf$index),]
names(mapdf) <- c('marker','chr','gmapdf','pmapdf')
#mapdf$duplicates <- duplicated(mapdf[c(2:3)])
mapdf <- mapdf[!duplicated(mapdf[c(2:3)]),]
#names(mapdf) <- c('marker','gmapdf','index','pmapdf','chr')

```

### Genome-wide scan
For each of the phenotype analyzed, permutations were used for each model to obtain genome-wide LOD significance threshold for p < 0.01, p < 0.05, p < 0.10,  respectively, separately for X and automsomes (A).  

The table shows the estimated significance thresholds from permutation test. 

We also looked at the kinship to see how correlated each sample is. Kinship values between pairs of samples range between 0 (no relationship) and 1.0 (completely identical). The darker the colour the more indentical the pairs are.

```{r permutation, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE, results="asis"}

Xcovar <- get_x_covar(gm)
#addcovar = model.matrix(~Sex, data = covars)[,-1]

#K <- calc_kinship(pr.qc, type = "loco")
#heatmap(K[[1]])
#K.overall <- calc_kinship(pr.qc, type = "overall")
#heatmap(K.overall)
kinship <- calc_kinship(pr.qc)
heatmap(kinship)

#operm <- scan1perm(pr.qc, gm$covar$phenos, Xcovar=Xcovar, n_perm=2000)
#operm <- scan1perm(pr.qc, gm$covar$phenos, addcovar = addcovar, n_perm=2000)
#operm <- scan1perm(pr.qc, gm$covar$phenos, n_perm=2000)
operm <- scan1perm(pr.qc, gm$covar["ICI.vs.EOI"], model="binary", n_perm=10, perm_Xsp=TRUE, chr_lengths=chr_lengths(gm$gmap))

summary_table<-data.frame(unclass(summary(operm, alpha=c(0.01,  0.05, 0.1))))
names(summary_table) <- c("autosomes","X")
summary_table$significance.level <- rownames(summary_table)

rownames(summary_table) <- NULL

summary_table[c(3,1:2)] %>%
  kable(escape = F,align = c("ccc")) %>%
  kable_styling("striped", full_width = T) %>%
  column_spec(1, bold=TRUE)

```
The figures below show QTL maps for each phenotype

```{r Genome-wide scan, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE, results="asis"}

out <- scan1(pr.qc, gm$covar["ICI.vs.EOI"], Xcovar=Xcovar, model="binary")

summary_table<-data.frame(unclass(summary(operm, alpha=c(0.01,  0.05, 0.1))))

plot_lod<-function(out,map){
  for (i in 1:dim(out)[2]){
    #png(filename=paste0("/Users/chenm/Documents/qtl/Jai/",colnames(out)[i],  "_lod.png"))
    
    #par(mar=c(5.1, 6.1, 1.1, 1.1))
    ymx <- maxlod(out) # overall maximum LOD score
    plot(out, gm$gmap, lodcolumn=i, col="slateblue", ylim=c(0, ymx+0.5))
    legend("topright", lwd=2, colnames(out)[i], bg="gray90")
    title(main = colnames(out)[i])
    #for (j in 1: dim(summary_table)[1]){
    #  abline(h=summary_table[j, i],col="red")
    #  text(x=400, y =summary_table[j, i]+0.12, labels = paste("p=", row.names(summary_table)[j]))
    #}
    #dev.off()
  }
}

plot_lod(out,gm$gmap)

```

### LOD peaks

The table below shows QTL peaks associated with the phenotype. We use the 95% threshold from the permutations to find peaks. 

```{r LOD peaks, fig.height = 6, fig.width = 9.5, fig.align = "center", echo=TRUE, results="asis"}

peaks<-find_peaks(out, gm$gmap, threshold=3.3, drop=1.5)

rownames(peaks) <- NULL
peaks[] %>%
  kable(escape = F,align = c("ccccccc")) %>%
  kable_styling("striped", full_width = T) %>%
  column_spec(1, bold=TRUE)

peaks_mbl <- list()
#corresponding info in Mb
for(i in 1:nrow(peaks)){
  lodindex <- peaks$lodindex[i]
  lodcolumn <- peaks$lodcolumn[i]
  chr <- peaks$chr[i]
  lod <- peaks$lod[i]
  pos <- mapdf$pmapdf[which(mapdf$gmapdf == peaks$pos[i] & mapdf$chr == peaks$chr[i])]
  ci_lo <- mapdf$pmapdf[which(mapdf$gmapdf == peaks$ci_lo[i] & mapdf$chr == peaks$chr[i])]
  ci_hi <- mapdf$pmapdf[which(mapdf$gmapdf == peaks$ci_hi[i] & mapdf$chr == peaks$chr[i])]
  peaks_mb=cbind(lodindex, lodcolumn, chr, pos, lod, ci_lo, ci_hi)
  peaks_mbl[[i]] <- peaks_mb
}
peaks_mba <- do.call(rbind, peaks_mbl)
peaks_mba <- as.data.frame(peaks_mba)
peaks_mba[,c("chr", "pos", "lod", "ci_lo", "ci_hi")] <- sapply(peaks_mba[,c("chr", "pos", "lod", "ci_lo", "ci_hi")], as.numeric)

rownames(peaks_mba) <- NULL
peaks_mba[] %>%
  kable(escape = F,align = c("ccccccc")) %>%
  kable_styling("striped", full_width = T) %>%
  column_spec(1, bold=TRUE)

```
### QTL effects

For each peak LOD location we give a list of gene

```{r QTL effects, echo=TRUE, results="asis"}

query_variants <- create_variant_query_func("/Users/corneb/Documents/MyJax/CS/Projects/support.files/qtl2/cc_variants.sqlite")
query_genes <- create_gene_query_func("/Users/corneb/Documents/MyJax/CS/Projects/support.files/qtl2/mouse_genes_mgi.sqlite")

for (i in 1:nrow(peaks)){
#for (i in 1:1){
  #Plot 1
  g <- maxmarg(pr.qc, gm$gmap, chr=peaks$chr[i], pos=peaks$pos[i], return_char=TRUE)
  #png(filename=paste0("/Users/chenm/Documents/qtl/Jai/","qtl_effect_", i, ".png"))
  #par(mar=c(4.1, 4.1, 1.5, 0.6))
  plot_pxg(g, gm$covar[,peaks$lodcolumn[i]], ylab=peaks$lodcolumn[i], sort=FALSE)
  title(main = paste("chr", chr=peaks$chr[i], "pos", peaks$pos[i], "(",peaks$lodcolumn[i],")"), line=0.2)
  ##dev.off()

  #chr = peaks$chr[i]

# Plot 2
  #coeff <- scan1coef(pr[,chr], cross$pheno[,peaks$lodcolumn[i]], addcovar = addcovar)
  #coeff <- scan1coef(pr[,chr], cross$pheno[,peaks$lodcolumn[i]], Xcovar=Xcovar)
  coeff <- scan1coef(pr.qc[,chr], gm$covar[peaks$lodcolumn[i]], model="binary")
  blup <- scan1blup(pr.qc[,chr], gm$covar[peaks$lodcolumn[i]])
  plot_coef(coeff, 
       gm$gmap[chr], columns=1:2,
       bgcolor="gray95", legend="bottomleft", 
       main = paste("chr", chr=peaks$chr[i], "pos", peaks$pos[i], "(",peaks$lodcolumn[i]," [scan1coeff])")
       )

  plot_coef(blup, 
       gm$gmap[chr], columns=1:2,
       bgcolor="gray95", legend="bottomleft", 
       main = paste("chr", chr=peaks$chr[i], "pos", peaks$pos[i], "(",peaks$lodcolumn[i]," [scan1blup])")
       )
  #last_coef <- unclass(coeff)[nrow(coeff),1:3]
  #for(t in seq(along=last_coef))
  #axis(side=4, at=last_coef[t], names(last_coef)[t], tick=FALSE)


 # Plot 3
  #c2effB <- scan1coef(pr.qc[,chr], gm$covar[peaks$lodcolumn[i]], model="binary", contrasts=cbind(a=c(-1, 0), d=c(0, -1)))
  #c2effBb <- scan1blup(pr.qc[,chr], gm$covar[peaks$lodcolumn[i]], contrasts=cbind(a=c(-1, 0), d=c(0, -1)))
  ##c2effB <- scan1coef(pr[,chr], cross$pheno[,peaks$lodcolumn[i]], addcovar = addcovar, contrasts=cbind(mu=c(1,1,1), a=c(-1, 0, 1), d=c(0, 1, 0)))
  ##c2effB <- scan1coef(pr[,chr], cross$pheno[,peaks$lodcolumn[i]],Xcovar=Xcovar, contrasts=cbind(mu=c(1,1,1), a=c(-1, 0, 1), d=c(0, 1, 0)))
  #plot(c2effB, gm$gmap[chr], columns=1:2,
  #     bgcolor="gray95", legend="bottomleft", 
  #     main = paste("chr", chr=peaks$chr[i], "pos", peaks$pos[i], "(",peaks$lodcolumn[i],")")
  #     )
  #plot(c2effBb, gm$gmap[chr], columns=1:2,
  #     bgcolor="gray95", legend="bottomleft", 
  #     main = paste("chr", chr=peaks$chr[i], "pos", peaks$pos[i], "(",peaks$lodcolumn[i],")")
  #     )
  ##last_coef <- unclass(c2effB)[nrow(c2effB),2:3] # last two coefficients
  ##for(t in seq(along=last_coef))
  ##  axis(side=4, at=last_coef[t], names(last_coef)[t], tick=FALSE)


  #Table 1
  chr = peaks_mba$chr[i]
  start=peaks_mba$ci_lo[i]
  end=peaks_mba$ci_hi[i]

  genesgss = query_genes(chr, start, end)

  rownames(genesgss) <- NULL
  genesgss$strand_old = genesgss$strand
  genesgss$strand[genesgss$strand=="+"] <- "positive"
  genesgss$strand[genesgss$strand=="-"] <- "negative"

  #genesgss <- 
  #table <- 
  #genesgss[,c("chr","type","start","stop","strand","ID","Name","Dbxref","gene_id","mgi_type","description")] %>%
  #kable(escape = F,align = c("ccccccccccc")) %>%
  #kable_styling("striped", full_width = T) #%>% 
  #cat #%>%
  #column_spec(1, bold=TRUE)
#
  #print(kable(genesgss[,c("chr","type","start","stop","strand","ID","Name","Dbxref","gene_id","mgi_type","description")], escape = F,align = c("ccccccccccc")))

  print(kable(genesgss[,c("chr","type","start","stop","strand","ID","Name","Dbxref","gene_id","mgi_type","description")], "html") %>% kable_styling("striped", full_width = T))

  #table
  

}

```