---
title: "Power Analysis - Myocarditis Yes vs Myocarditis No"
author: "Belinda Cornes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)

library("kableExtra")
library("knitr")
options(stringsAsFactors = FALSE)
library(data.table) 
library(tidyr)     
library(mclust)     
library(rhdf5)      
library(optparse)
library(dplyr)
library(cluster)
library(readxl)
library(psych)
library(tibble)
library(ggplot2)
library(reshape2)
library(qtl2)
#library(qtl)
library(abind)
#install.packages("qtlDesign")
library(qtlDesign)

setwd("/projects/serreze-lab/USERS/corneb/qtl2/workflowr/2023.04.13/Serreze-T1D_Workflow")

```

## Loading Data

```{r data, echo=TRUE}

load("data/gm_allqc_7.batches_myo_mis.RData")

#gm_allqc
gm=gm_allqc
gm

#pr <- readRDS("data/serreze_probs_allqc.rds")
#pr <- readRDS("data/serreze_probs.rds")

##extracting animals with ici and pbs group status
##extracting animals 
#Myocarditis Status
miceinfo <- gm$covar[gm$covar$"Myocarditis Status" == "YES" | gm$covar$"Myocarditis Status" == "NO",]
#table(miceinfo$group)
#MMurine MHC KO Status
miceinfo <- miceinfo[miceinfo$"Murine MHC KO Status" == "HOM",]
#Drug Treatment
#miceinfo <- miceinfo[miceinfo$"Drug Treatment" == "PBS",]
#Clinical Phenotype
miceinfo <- miceinfo[miceinfo$"clinical pheno" == "SICK" | miceinfo$"clinical pheno" == "EOI",]

table(miceinfo$"Myocarditis Status")
table(miceinfo$"Murine MHC KO Status")
table(miceinfo$"Drug Treatment")
table(miceinfo$"clinical pheno")

mice.ids <- rownames(miceinfo)

##removing X samples
x.samples = c("NM00179-ICI-EOI","NM00186-ICI-EOI","D22-ICI-EOI","D39-ICI-Myo","D67-ICI-Myo","D111-PBS-Myo","D120-ICI-Myo","6895-PBS-Myo","D704-ICI-Myo")
mice.ids = mice.ids[!as.character(do.call(rbind.data.frame, strsplit(mice.ids, "_"))[,7]) %in% x.samples]

#removing duplicates
dup.samples = c("The_Jackson_Lab_Serreze_MURGIGV01_20230225_D249-PBS-EOI_H3")
mice.ids = mice.ids[!mice.ids %in% dup.samples]

gm <- gm[mice.ids]
gm
table(gm$covar$"Myocarditis Status")
table(gm$covar$"Murine MHC KO Status")
table(gm$covar$"Drug Treatment")
table(gm$covar$"clinical pheno")

gm$covar$myo.yes_vs_myo.no <- ifelse(gm$covar$"Myocarditis Status" == "NO", 0, 1)
gm$covar$group = gm$covar$myo.yes_vs_myo.no


##loading genotypes

genos.1 <- read.csv("data/myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_7.batches_myo_mis.csv")
dim(genos.1)
genos.1 <- genos.1[genos.1$Include,]
dim(genos.1)
#dis.genos.1 <- read.csv("data/myo-yes.vs.myo-no_marker.freq_low.geno.freq.removed_geno.ratio_7.batches_myo_mis_mis.csv")
#dim(dis.genos.1)
rownames(genos.1) <- genos.1$marker
genos.1 <- t(genos.1[3:(ncol(genos.1)-9)])
rownames(genos.1) <- gsub("\\.","-",rownames(genos.1))

genos.2<- read.csv("data/myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_sample.outliers.removed_7.batches_myo_mis.csv")
dim(genos.2)
genos.2 <- genos.2[genos.2$Include,]
dim(genos.2)
#dis.genos.2 <- read.csv("data/myo-yes.vs.myo-no_marker.freq_low.geno.freq.removed_sample.outliers.removed_geno.ratio_7.batches_myo_mis_mis.csv")
rownames(genos.2) <- genos.2$marker
genos.2 <- t(genos.2[3:(ncol(genos.2)-9)])
rownames(genos.2) <- gsub("\\.","-",rownames(genos.2))


##merge with phenotypes (age.of.onset & binary)

names(gm$covar)[6] <- c("age.of.onset")
gm$covar$age.of.onset <- as.numeric(gm$covar$age.of.onset)
#if(0 %in% unique(gm$covar$age.of.onset)){
gm$covar[gm$covar$age.of.onset==0 | is.na(gm$covar$age.of.onset) | gm$covar$age.of.onset=="NA" ,]$age.of.onset = ""
gm$covar[gm$covar$age.of.onset=='' ,]$age.of.onset <- 40
gm$covar$age.of.onset <- as.numeric(gm$covar$age.of.onset)

geno.pheno.1 <- merge(data.frame(genos.1), gm$covar,  by="row.names", all.x=T, sort=F)
aov.1 <- stats::aov(age.of.onset ~ group, geno.pheno.1)
mse.1 <- mean(aov.1$residuals^2) 
geno.pheno.2 <- merge(data.frame(genos.2), gm$covar,  by="row.names", all.x=T, sort=F)
aov.2 <- stats::aov(age.of.onset ~ group, geno.pheno.2)
mse.2 <- mean(aov.2$residuals^2) 


## calculating effect (age.of.onset)

age.marker.sum.1 <- list()
geno.pheno.1a <- geno.pheno.1[-1]
for(i in 1:(ncol(geno.pheno.1a)-13)){
#geno.pheno.1$snp=geno.pheno.1[i]
age.marker.1 <- geno.pheno.1a %>% 
  group_by(geno.pheno.1a[i])  %>%
  summarize(mean_age.of.onset = mean(as.numeric(age.of.onset), na.rm =T))


  names(age.marker.1) <- c("genotypes", paste0(names(geno.pheno.1a[i]),"_age.of.onset"))
  genotypes <- as.data.frame(c("-","AA","AB"))
  names(genotypes) <- c("genotypes")
  age.marker.1 <- merge(genotypes, age.marker.1, by=c("genotypes"), all=T, sort=T)
  #print(age.marker.1)

  age.marker.sum.1[[i]] <- as.data.frame(age.marker.1)
}

age.marker.sum.1.all <- Reduce(function(x, y) merge(x, y, by=c("genotypes"),all=TRUE),age.marker.sum.1)
rownames(age.marker.sum.1.all) <- age.marker.sum.1.all$genotypes
age.marker.sum.1.all.t <- t(age.marker.sum.1.all[-1])
rownames(age.marker.sum.1.all.t) <- gsub("_age.of.onset","",rownames(age.marker.sum.1.all.t))
age.marker.sum.1.all.t <- as.data.frame(age.marker.sum.1.all.t)
age.marker.sum.1.all.t$diffhethom <- age.marker.sum.1.all.t$AB - age.marker.sum.1.all.t$AA
max.1 <- max(age.marker.sum.1.all.t$diffhethom)
min.1 <- min(age.marker.sum.1.all.t$diffhethom)
write.csv(age.marker.sum.1.all.t,"data/mean.differences_age.of.onset_myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_7.batches_myo_mis.csv")

age.marker.sum.2 <- list()
geno.pheno.2a <- geno.pheno.2[-1]
for(i in 1:(ncol(geno.pheno.2a)-13)){
#geno.pheno.2$snp=geno.pheno.2[i]
age.marker.2 <- geno.pheno.2a %>% 
  group_by(geno.pheno.2a[i])  %>%
  summarize(mean_age.of.onset = mean(as.numeric(age.of.onset), na.rm =T))


  names(age.marker.2) <- c("genotypes", paste0(names(geno.pheno.2a[i]),"_age.of.onset"))
  genotypes <- as.data.frame(c("-","AA","AB"))
  names(genotypes) <- c("genotypes")
  age.marker.2 <- merge(genotypes, age.marker.2, by=c("genotypes"), all=T, sort=T)
  #print(age.marker.2)

  age.marker.sum.2[[i]] <- as.data.frame(age.marker.2)
}

age.marker.sum.2.all <- Reduce(function(x, y) merge(x, y, by=c("genotypes"),all=TRUE),age.marker.sum.2)
rownames(age.marker.sum.2.all) <- age.marker.sum.2.all$genotypes
age.marker.sum.2.all.t <- t(age.marker.sum.2.all[-1])
rownames(age.marker.sum.2.all.t) <- gsub("_age.of.onset","",rownames(age.marker.sum.2.all.t))
age.marker.sum.2.all.t <- as.data.frame(age.marker.sum.2.all.t)
age.marker.sum.2.all.t$diffhethom <- age.marker.sum.2.all.t$AB - age.marker.sum.2.all.t$AA
max.2 <- max(age.marker.sum.2.all.t$diffhethom)
min.2 <- min(age.marker.sum.2.all.t$diffhethom)
write.csv(age.marker.sum.2.all.t,"data/mean.differences_age.of.onset_myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_sample.outliers.removed_7.batches_myo_mis.csv")


max.a <- max(max.1, max.2)
print(max.a)

min.a <- min(min.1, min.2)
print(min.a)

mse.a <- (mse.1 + mse.2)/2
print(mse.a)


## calculating effect (group)

age.marker.sum.1 <- list()
geno.pheno.1a <- geno.pheno.1[-1]
for(i in 1:(ncol(geno.pheno.1a)-13)){
#geno.pheno.1$snp=geno.pheno.1[i]
age.marker.1 <- geno.pheno.1a %>% 
  group_by(geno.pheno.1a[i])  %>%
  summarize(mean_group = mean(as.numeric(group), na.rm =T))


  names(age.marker.1) <- c("genotypes", paste0(names(geno.pheno.1a[i]),"_group"))
  genotypes <- as.data.frame(c("-","AA","AB"))
  names(genotypes) <- c("genotypes")
  age.marker.1 <- merge(genotypes, age.marker.1, by=c("genotypes"), all=T, sort=T)
  #print(age.marker.1)

  age.marker.sum.1[[i]] <- as.data.frame(age.marker.1)
}

age.marker.sum.1.all <- Reduce(function(x, y) merge(x, y, by=c("genotypes"),all=TRUE),age.marker.sum.1)
rownames(age.marker.sum.1.all) <- age.marker.sum.1.all$genotypes
age.marker.sum.1.all.t <- t(age.marker.sum.1.all[-1])
rownames(age.marker.sum.1.all.t) <- gsub("_group","",rownames(age.marker.sum.1.all.t))
age.marker.sum.1.all.t <- as.data.frame(age.marker.sum.1.all.t)
age.marker.sum.1.all.t$diffhethom <- age.marker.sum.1.all.t$AB - age.marker.sum.1.all.t$AA
write.csv(age.marker.sum.1.all.t,"data/mean.differences_group_myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_7.batches_myo_mis.csv")

age.marker.sum.2 <- list()
geno.pheno.2a <- geno.pheno.2[-1]
for(i in 1:(ncol(geno.pheno.2a)-13)){
#geno.pheno.2$snp=geno.pheno.2[i]
age.marker.2 <- geno.pheno.2a %>% 
  group_by(geno.pheno.2a[i])  %>%
  summarize(mean_group = mean(as.numeric(group), na.rm =T))


  names(age.marker.2) <- c("genotypes", paste0(names(geno.pheno.2a[i]),"_group"))
  genotypes <- as.data.frame(c("-","AA","AB"))
  names(genotypes) <- c("genotypes")
  age.marker.2 <- merge(genotypes, age.marker.2, by=c("genotypes"), all=T, sort=T)
  #print(age.marker.2)

  age.marker.sum.2[[i]] <- as.data.frame(age.marker.2)
}

age.marker.sum.2.all <- Reduce(function(x, y) merge(x, y, by=c("genotypes"),all=TRUE),age.marker.sum.2)
rownames(age.marker.sum.2.all) <- age.marker.sum.2.all$genotypes
age.marker.sum.2.all.t <- t(age.marker.sum.2.all[-1])
rownames(age.marker.sum.2.all.t) <- gsub("_group","",rownames(age.marker.sum.2.all.t))
age.marker.sum.2.all.t <- as.data.frame(age.marker.sum.2.all.t)
age.marker.sum.2.all.t$diffhethom <- age.marker.sum.2.all.t$AB - age.marker.sum.2.all.t$AA
write.csv(age.marker.sum.2.all.t,"data/mean.differences_group_myo-yes.vs.myo-no_sample.genos_marker.freq_low.geno.freq.removed_sample.outliers.removed_7.batches_myo_mis.csv")


max.g <- max(max.1, max.2)
print(max.g)

min.g <- min(min.1, min.2)
print(min.g)

mse.g <- (mse.1 + mse.2)/2
print(mse.g)

```


## Power

For powercalc the power is returned, along with the proportion of variance explained.  LOD threshold set to 3 (which is roughly what is seen for suggestive significance)

```{r power, echo=TRUE}


## age.of onset


powercalc(cross = "bc",n=length(mice.ids),effect=min.a,sigma2=mse.a,thresh=3,sel.frac=1,theta=0,bio.reps=1)
powercalc(cross = "bc",n=length(mice.ids),effect=max.a,sigma2=mse.a,thresh=3,sel.frac=1,theta=0,bio.reps=1)

#powercalc("bc",100,31,sigma2=1,sel.frac=1,theta=0)

## group (binary)

powercalc(cross = "bc",n=length(mice.ids),effect=min.g,sigma2=mse.g,thresh=3,sel.frac=1,theta=0,bio.reps=1)
powercalc(cross = "bc",n=length(mice.ids),effect=max.g,sigma2=mse.g,thresh=3,sel.frac=1,theta=0,bio.reps=1)

## graph

```

## Sample Size (with 80% power)

For samplesize the sample size (rounded up to the nearest integer) is returned along with the proportion of variance explained. LOD threshold set to 3 (which is roughly what is seen for suggestive significance) and 80% power

```{r sample, echo=TRUE}


## age.of onset
samplesize(cross = "bc",effect=min.a,sigma2=mse.a,thresh=3,sel.frac=1,theta=0,bio.reps=1)
samplesize(cross = "bc",effect=max.a,sigma2=mse.a,thresh=3,sel.frac=1,theta=0,bio.reps=1)

## group (binary)
samplesize(cross = "bc",effect=min.g,sigma2=mse.g,thresh=3,sel.frac=1,theta=0,bio.reps=1)
samplesize(cross = "bc",effect=max.g,sigma2=mse.g,thresh=3,sel.frac=1,theta=0,bio.reps=1)

## graph

```